### 2023-02-07 report/results.Rmd
Upload mixture model tutorial

---
title: "Modelling writing hesitations in text writing as finite mixture process"
author: "Jens Roeser"
date: "Compiled `r format(Sys.Date(), '%b %d %Y')`"
output: 
  rmdformats::downcute:
    keep_md: true
    self_contained: true
    thumbnails: false # for images
    lightbox: true
    gallery: false
    highlight: tango
    use_bookdown: true # for cross references
bibliography: references.bib
csl: apa.csl
link-citations: yes
---

```{r include = F}
library(tidyverse)
library(kableExtra)
library(janitor)
library(brms)
library(polspline)
library(patchwork)
library(rmdformats)
library(gridExtra)
library(grid)
library(gtable)
library(gridExtra)
library(scales)
#library(caret)
#library(InformationValue)

# Calculate Bayes Factor
BF <- function(ps, prior_sd = 1){
  fit_posterior <- logspline(ps) 
  posterior <- dlogspline(0, fit_posterior) # Height of the posterior at 0 
  prior <- dnorm(0, 0, prior_sd) # Height of the prior at 0
  BF10 <- prior / posterior 
  return(BF10)
}

logit <- function(p) log(p / (1-p))
ilogit <- function(x) 1 / (1 + exp(-x))
inv_logit <- function(logit) exp(logit) / (1 + exp(logit))

se_bin <- function(x) sqrt((mean(x, na.rm = T)*(1 - mean(x, na.rm = T)))/length(x)) # se for binary data

lower <- function(x) quantile(x, prob = .025)
upper <- function(x) quantile(x, prob = .975)

# Remove leading zeros and round numbers
dezero <- function(x, dp){ # dp is decimal places
  fmt = paste0('%.',dp,'f')
  x = sprintf(fmt,x)
  x = str_replace(x, '(-|^)0\\.','\\1\\.')
  return(x)
}
dezero_plot <- function(x) dezero(x, 1)

theme_set(theme_bw(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

options(scipen = 999)

beta_label <- "Short key transition\nduration"
delta_label <- "Slowdown for hesitant key\ntransitions"
theta_label <- "Probability of hesitant\ntransitions"
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      comment = NA,
                      #                      dev = c("pdf", "png"), 
                      #                     fig.keep = "all", 
                      #                    fig.path = "figures/",
                      fig.align = 'center', 
                      #                   pdf.options(encoding = "ISOLatin9.enc"),
                      fig.width = 8, 
                      fig.height = 5,
                      width=90)
```

# Analysis

We reanalysed data sets including process information from participants writing text. For all data sets we fit a series of four models each with random effects for participants. Probability functions used were normal and log-normal in line with typically treatments used in the literature, a log-normal distribution with unequal variances for model predictions, and a bimodal mixed effects model. Stan code for mixture models was based on @roeser2021modelling. Text locations (levels: before sentence, before word, within word) was included as predictor in all models.

Data were analysed in Bayesian mixed effects models [@gelman2014;@mcelreath2016statistical]. The R [@R-base] package rstan [@rstan] was used to interface with the probabilistic programming language Stan [@carpenter2016stan] which was used to implement all models. Models were fitted with weakly informative priors [see @mcelreath2016statistical], and run with 10,000 iterations on 3 chains with a warm-up of 5,000 iterations and no thinning. Model convergence was confirmed by the Rubin-Gelman statistic ($\hat{R}$ = 1) [@gelman1992] and inspection of the Markov chain Monte Carlo chains.

For model comparisons we used out-of-sample predictions estimated using Pareto smoothed importance-sampling leave-one-out cross-validation [@vehtari2015pareto; @vehtari2017practical]. Predictive performance was estimated as the sum of the expected log predictive density ($\widehat{elpd}$) and the difference $\Delta\widehat{elpd}$ between models. The advantage of using leave-one-out cross-validation is that models with more parameters are penalised to prevent overfit.

# Data processing

TODO: Add a description with general data filtering and a table with how many data were removed from the analysis by study.

# C2L1

The C2L1 data set comprises data Norwegian 6th graders--*N*=133, mean age 11 years 10 months--published in @ronneberg2022process. The children composed argumentative essays in Norwegian, a language with a relatively shallow orthography.

```{r eval = F}
d <- read_csv("../data/c2l1.csv") %>%
  filter(!is.na(iki), 
         iki > 50, 
         iki < 30000) %>%
  mutate(ppt = as.numeric(factor(subno)),
         condition = factor(str_c(location, sep = "_")),
         cond_num = as.integer(condition)) %>% 
  select(ppt, iki, condition, cond_num)

count(d, ppt) %>% 
  filter(n < 100)
pull(d, ppt) %>% unique() %>% length()
```

TODO: might need to remove kids that don't speak Norwegian at home (see github issue).

## Model comparisons

### Fit to data

```{r predictionsc2l1, fig.cap="C2L1 data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

files <- list.files("../stanout/c2l1", pattern = "sims", full.names = T)
sims <- read_csv(files) %>% 
  pivot_longer(cols = c(y_obs, y_tilde)) %>% 
  mutate(across(name, recode_factor, 
                y_tilde = "Simulated data",
                y_obs = "Observed data",
                .ordered = T)) %>% 
  filter(name ==  "Simulated data" |
         (name == "Observed data" & sim_idx == 1))

max_iki <- 2000

plot_mog <- sims %>%
  filter(model == "mog", value < max_iki) %>% 
  mutate(across(model, recode, mog = "Bimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmm <- sims %>%
  filter(model == "lmm", value < max_iki) %>% 
  mutate(across(model, recode, lmm = "Unimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmgaus <- sims %>%
  filter(model == "lmmgaus", value < max_iki) %>% 
  mutate(across(model, recode, lmmgaus = "Unimodal normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmuneqvar <- sims %>%
  filter(model == "lmmuneqvar", value < max_iki) %>% 
  mutate(across(model, recode, lmmuneqvar = "Unimodal log-normal (unequal variance)")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plots <- plot_lmmgaus + plot_lmm + plot_lmmuneqvar + plot_mog +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Out-of-samples cross-validation

```{r}
files <- list.files("../stanout/c2l1", pattern = "^modelcomp", full.names = T)
table <- purrr::map_dfr(files, read_csv) %>% 
  mutate(across(location, replace_na, "overall")) %>% 
  select(location, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(across(location, ~str_c("Text location: ",.)),
            across(model, recode, lmm = "Unimodal log-normal",
                                  lmmgaus = "Unimodal normal",
                                  mog = "Bimodal log-normal",
                                  lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = location, values_from = contains("elpd")) %>% 
  select(Model = model, 
         ends_with("overall"), 
         ends_with("sent"), 
         ends_with("beforeword"), 
         ends_with("withinword"))

```

```{r c2l1, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"),4)

kable(table, caption = "C2L1 data. Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 8))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "All transitions" = 2,
                     "Before sentence" = 2, 
                     "Before word" = 2,
                     "Within word" = 2)) %>%
  column_spec(1, width = "8em")

```

## Posterior parameter estimates of mixture model

```{r c2l1post, fig.cap="C2l1. Posterior parameter distribution"}
file <- list.files(path = "../stanout/c2l1", pattern = "mog.csv$",full.names = T)
ps <- read_csv(file) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(location, name) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"))

beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = "",
       shape = "") +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

beta + delta + theta +
  plot_layout(guides = "collect")

```

# CATO

Data are published in @torrance2016adolescent. Norwegian upper
secondary students--*N*=26, mean age = 16.9 years--with weak decoding skills and 26 age-matched controls composed expository texts by keyboard under two conditions: normally and with letters masked to prevent them reading what they were writing.

## Model comparisons

### Fit to data

```{r predictionscato, fig.cap="CATO data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

files <- list.files("../stanout/cato", pattern = "sims", full.names = T)
sims <- read_csv(files) %>% 
  pivot_longer(cols = c(y_obs, y_tilde)) %>% 
  mutate(across(name, recode_factor, 
                y_tilde = "Simulated data",
                y_obs = "Observed data",
                .ordered = T)) %>% 
  filter(name ==  "Simulated data" |
         (name == "Observed data" & sim_idx == 1))

max_iki <- 2000

plot_mog <- sims %>%
  filter(model == "mog", value < max_iki) %>% 
  mutate(across(model, recode, mog = "Bimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmm <- sims %>%
  filter(model == "lmm", value < max_iki) %>% 
  mutate(across(model, recode, lmm = "Unimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmgaus <- sims %>%
  filter(model == "lmmgaus", value < max_iki) %>% 
  mutate(across(model, recode, lmmgaus = "Unimodal normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmuneqvar <- sims %>%
  filter(model == "lmmuneqvar", value < max_iki) %>% 
  mutate(across(model, recode, lmmuneqvar = "Unimodal log-normal (unequal variance)")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plots <- plot_lmmgaus + plot_lmm + plot_lmmuneqvar + plot_mog +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Out-of-samples cross-validation

```{r}
files <- list.files("../stanout/cato", pattern = "^modelcomp", full.names = T)
table <- purrr::map_dfr(files, read_csv) %>% 
  mutate(across(location, replace_na, "overall")) %>% 
  select(location, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(across(location, ~str_c("Text location: ",.)),
            across(model, recode, lmm = "Unimodal log-normal",
                                  lmmgaus = "Unimodal normal",
                                  mog = "Bimodal log-normal",
                                  lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = location, values_from = contains("elpd")) %>% 
  select(Model = model, 
         ends_with("overall"), 
         ends_with("sent"), 
         ends_with("beforeword"), 
         ends_with("withinword"))

```

```{r cato, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"),4)

kable(table, caption = "CATO data. Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 8))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "All transitions" = 2,
                     "Before sentence" = 2, 
                     "Before word" = 2,
                     "Within word" = 2)) %>%
  column_spec(1, width = "8em")

```

## Posterior parameter estimates of mixture model

```{r catopost, fig.cap="CATO. Posterior parameter distribution"}
file <- list.files(path = "../stanout/cato", pattern = "mog.csv$",full.names = T)
ps <- read_csv(file) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(task, group, location, name) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}")) %>% 
  mutate(task = str_c("Task: ", task, "; Group: ", group))

beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = "",
       shape = "") +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = 1:4) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

beta + delta + theta +
  plot_layout(guides = "collect")

```

# SPL2

Data are going to be published in @torrance.

## Methods

Undergraduate university students--*N* = 39, 28 female, mean age = 20.6 years (SD = 1.51)--wrote two short argumentative essays, one in English (the student's first language in all cases; L1) and one in Spanish (L2) using CyWrite [@chukharev2019combined]. CyWrite provides a writing environment with basic word processing functionality (e.g., Microsoft WordPad), including text selection by mouse action, and copy-and-paste. We recorded the time of each keystroke and mouse action, and tracked writers' eye movements within their emerging text.

Writing tasks: Participants were given a 40 minute time limit. They wrote essays in response to each of two prompts, with order and L1 / L2 counterbalanced across subjects.

## Model comparisons

### Fit to data

```{r predictionsspl2, fig.cap="SPL2 data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

files <- list.files("../stanout/spl2", pattern = "sims", full.names = T)
sims <- read_csv(files) %>% 
  pivot_longer(cols = c(y_obs, y_tilde))  %>% 
  mutate(across(name, recode_factor, 
                y_tilde = "Simulated data",
                y_obs = "Observed data",
                .ordered = T)) %>% 
  filter(name ==  "Simulated data" |
         (name == "Observed data" & sim_idx == 1))

max_iki <- 2000

plot_mog <- sims %>%
  filter(model == "mog", value < max_iki) %>% 
  mutate(across(model, recode, mog = "Bimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmm <- sims %>%
  filter(model == "lmm", value < max_iki) %>% 
  mutate(across(model, recode, lmm = "Unimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmgaus <- sims %>%
  filter(model == "lmmgaus", value < max_iki) %>% 
  mutate(across(model, recode, lmmgaus = "Unimodal normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmuneqvar <- sims %>%
  filter(model == "lmmuneqvar", value < max_iki) %>% 
  mutate(across(model, recode, lmmuneqvar = "Unimodal log-normal (unequal variance)")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plots <- plot_lmmgaus + plot_lmm + plot_lmmuneqvar + plot_mog +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Out-of-samples cross-validation

```{r}
files <- list.files("../stanout/spl2", pattern = "^modelcomp", full.names = T)
table <- purrr::map_dfr(files, read_csv) %>% 
  mutate(across(location, replace_na, "overall")) %>% 
  select(location, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(across(location, ~str_c("Text location: ",.)),
            across(model, recode, lmm = "Unimodal log-normal",
                                  lmmgaus = "Unimodal normal",
                                  mog = "Bimodal log-normal",
                                  lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = location, values_from = contains("elpd")) %>% 
  select(Model = model, 
         ends_with("overall"), 
         ends_with("sent"), 
         ends_with("beforeword"), 
         ends_with("withinword"))

```

```{r spl2, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"),4)

kable(table, caption = "SPL2 data. Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 8))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "All transitions" = 2,
                     "Before sentence" = 2, 
                     "Before word" = 2,
                     "Within word" = 2)) %>%
  column_spec(1, width = "8em")

```

## Posterior parameter estimates of mixture model

```{r spl2post, fig.cap="SPL2. Posterior parameter distribution"}
file <- list.files(path = "../stanout/spl2", pattern = "mog.csv$",full.names = T)
ps <- read_csv(file) %>% 
  mutate(across(lang, recode, "EN" = "L1", "ES" = "L2")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(lang, location, name) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"))

beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = lang,
             shape = lang)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = "Language",
       shape = "Language")

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = lang,
             shape = lang)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "Language",
       shape = "Language") +
  theme(axis.title.y = element_blank())

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = lang,
             shape = lang)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "Language",
       shape = "Language") +
  theme(axis.title.y = element_blank())

beta + delta + theta +
  plot_layout(guides = "collect")

```

# PLanTra

## Methods

The PLanTra (Plain Language for Financial Content: Assessing the Impact of Training on Students' Revisions and Readers' Comprehension) data set [@rossetti2022s] involved the collection of keystroke data from 47 university students, who were randomly divided into an experimental and a control group. In a pre-test session, all students were assigned an extract of a corporate report dealing with sustainability and were instructed to revise it to make it easier to read for a lay audience. Subsequently, the experimental group received training on how to apply plain language principles to sustainability content, while the control group received training exclusively on the topic of sustainability. During a post-test session, both groups were instructed to revise a second extract of a corporate sustainability report with the same goal--i.e. making it easier to read for a lay audience--by applying what they had learned from their respective training. The texts were in English while the participants were native speakers of other languages (mainly Dutch), so writing took place in second language. It should be pointed out that, while some students decided to revise the assigned texts, the majority of them opted for rewriting the texts from scratch.

## Model comparisons

### Fit to data

```{r predictionsplantra, fig.cap="PLanTra data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

files <- list.files("../stanout/plantra", pattern = "sims", full.names = T)
sims <- read_csv(files) %>% 
  pivot_longer(cols = c(y_obs, y_tilde))  %>% 
  mutate(across(name, recode_factor, 
                y_tilde = "Simulated data",
                y_obs = "Observed data",
                .ordered = T)) %>% 
  filter(name ==  "Simulated data" |
         (name == "Observed data" & sim_idx == 1))

max_iki <- 2000

plot_mog <- sims %>%
  filter(model == "mog", value < max_iki) %>% 
  mutate(across(model, recode, mog = "Bimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmm <- sims %>%
  filter(model == "lmm", value < max_iki) %>% 
  mutate(across(model, recode, lmm = "Unimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmgaus <- sims %>%
  filter(model == "lmmgaus", value < max_iki) %>% 
  mutate(across(model, recode, lmmgaus = "Unimodal normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmuneqvar <- sims %>%
  filter(model == "lmmuneqvar", value < max_iki) %>% 
  mutate(across(model, recode, lmmuneqvar = "Unimodal log-normal (unequal variance)")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plots <- plot_lmmgaus + plot_lmm + plot_lmmuneqvar + plot_mog +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Out-of-samples cross-validation

```{r}
files <- list.files("../stanout/plantra", pattern = "^modelcomp", full.names = T)
table <- purrr::map_dfr(files, read_csv) %>% 
  mutate(across(location, replace_na, "overall")) %>% 
  select(location, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(across(location, ~str_c("Text location: ",.)),
            across(model, recode, lmm = "Unimodal log-normal",
                                  lmmgaus = "Unimodal normal",
                                  mog = "Bimodal log-normal",
                                  lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = location, values_from = contains("elpd")) %>% 
  select(Model = model, 
         ends_with("overall"), 
         ends_with("sent"), 
         ends_with("beforeword"), 
         ends_with("withinword"))
```

```{r plantra, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"),4)

kable(table, caption = "PLanTra data. Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 8))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "All transitions" = 2,
                     "Before sentence" = 2, 
                     "Before word" = 2,
                     "Within word" = 2)) %>%
  column_spec(1, width = "8em")

```

## Posterior parameter estimates of mixture model

```{r plantrapost, fig.cap="PLanTra. Posterior parameter distribution"}
file <- list.files(path = "../stanout/plantra", pattern = "mog.csv$",full.names = T)
ps <- read_csv(file) %>% 
  mutate(across(task, factor, levels = c("pre-test", "post-test"), ordered = T)) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(task, location, name) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"))

beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = "Task",
       shape = "Task") +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "Task",
       shape = "Task") +
  theme(axis.title.y = element_blank()) +
    guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = task,
             shape = task)) +
  geom_point(size = 3, position = position_dodge(.25)) +
  geom_errorbar(width = 0, position = position_dodge(.25)) +
  facet_grid(~name) +
  scale_shape_manual(values = c(21, 23)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "Task",
       shape = "Task") +
  theme(axis.title.y = element_blank()) +
    guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

beta + delta + theta +
  plot_layout(guides = "collect")

```

# LIFT

## Methods

LIFT (Improving Pre-university Students' Performance in Academic Synthesis Tasks with Level-up Instructions and Feedback Tool) [@nina_vandermeulen_2020_3893538].

## Model comparisons

### Fit to data

```{r predictionslift, fig.cap="LIFT data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

files <- list.files("../stanout/lift", pattern = "sims", full.names = T)
sims <- read_csv(files) %>% 
  pivot_longer(cols = c(y_obs, y_tilde))  %>% 
  mutate(across(name, recode_factor, 
                y_tilde = "Simulated data",
                y_obs = "Observed data",
                .ordered = T)) %>% 
  filter(name ==  "Simulated data" |
         (name == "Observed data" & sim_idx == 1))

max_iki <- 2000

plot_mog <- sims %>%
  filter(model == "mog", value < max_iki) %>% 
  mutate(across(model, recode, mog = "Bimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmm <- sims %>%
  filter(model == "lmm", value < max_iki) %>% 
  mutate(across(model, recode, lmm = "Unimodal log-normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmgaus <- sims %>%
  filter(model == "lmmgaus", value < max_iki) %>% 
  mutate(across(model, recode, lmmgaus = "Unimodal normal")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plot_lmmuneqvar <- sims %>%
  filter(model == "lmmuneqvar", value < max_iki) %>% 
  mutate(across(model, recode, lmmuneqvar = "Unimodal log-normal (unequal variance)")) %>% 
  ggplot(aes(x = value, 
             group = interaction(sim_idx,name), 
             colour = name)) +
  geom_density() +
  facet_wrap(~model) +
  scale_colour_manual(values = c("firebrick3", "black")) +
  scale_x_continuous(labels = scales::comma) +
  labs(colour = "") +
  theme(axis.title = element_blank())

plots <- plot_lmmgaus + plot_lmm + plot_lmmuneqvar + plot_mog +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Out-of-samples cross-validation

```{r}
files <- list.files("../stanout/lift", pattern = "^modelcomp", full.names = T)
table <- purrr::map_dfr(files, read_csv) %>% 
  mutate(across(location, replace_na, "overall")) %>% 
  select(location, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(across(location, ~str_c("Text location: ",.)),
            across(model, recode, lmm = "Unimodal log-normal",
                                  lmmgaus = "Unimodal normal",
                                  mog = "Bimodal log-normal",
                                  lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = location, values_from = contains("elpd")) %>% 
  select(Model = model, 
         ends_with("overall"), 
         ends_with("sent"), 
         ends_with("beforeword"), 
         ends_with("withinword"))
```

```{r lift, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"),4)

kable(table, caption = "LIFT data. Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 8))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "All transitions" = 2,
                     "Before sentence" = 2, 
                     "Before word" = 2,
                     "Within word" = 2)) %>%
  column_spec(1, width = "8em")

```

## Posterior parameter estimates of mixture model

```{r}
file <- list.files(path = "../stanout/lift", pattern = "mog.csv$",full.names = T)
ps <- read_csv(file) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(genre, topic, location, name) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}")) %>% 
  ungroup() %>% 
  mutate(group = str_c("Topic: ", topic, "; Genre: ", genre))
```

```{r liftpost, fig.cap="LIFT. Posterior parameter distribution"}
posd <- position_dodge(.65)
dotsize <- 2.5
beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = group,
             shape = group)) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1:8)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = "",
       shape = "")

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = group,
             shape = group)) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1:8)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = group,
             shape = group)) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1:8)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

beta + delta + theta +
  plot_layout(guides = "collect")

```

# Cross-task / data set comparisons

```{r}
# Get the differences for tasks / across data set
# use effect size = beta / sd
# averaged across grouping variable

# Also look at by ppt estimates?

```

```{r}
files <- str_c("../stanout/", 
               c("lift", "spl2", "plantra", "cato", "c2l1"), 
               "/mog.csv")
ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, str_remove_all, 
                "\\.|\\/|stanout|\\mog.csv"),
         across(data, recode_factor, 
                "cato" = "CATO (non-dyslexic\nunmasked)",
                "spl2" = "SPL2 (L1)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                .ordered = TRUE)) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked")) %>% 
#  filter(str_detect(data, "CATO"))
  pivot_wider(names_from = param, 
              values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(delta, ~.+beta),
         across(c(delta, beta), exp),
         across(delta, ~.-beta),
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:prob) %>% 
  group_by(data, location, name) %>% 
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper),
                   .names = "{.fn}")) %>% 
  ungroup()
```

```{r crossstudypost, fig.cap="Across studies. Posterior parameter distribution"}
posd <- position_dodge(.65)
dotsize <- 2.5
grouplabel <- "Data set"
beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = data,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = data,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             colour = data,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_point(size = dotsize, position = posd) +
  geom_errorbar(width = 0, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9)) +
  scale_colour_viridis_d(end = .7) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

beta + delta + theta +
  plot_layout(guides = "collect")

```

```{r eval = F}
# use posterior to calculate interactions, cell means and dataset differences
files <- str_c("../stanout/", c("lift", "spl2", "plantra", "cato", "c2l1"), "/mog.csv")

plantra <- read_csv(files[1], id = "data") %>% 
  pivot_wider(names_from = c(genre, topic), values_from = value) %>% 
  unnest(cols = c(ARG_A, ARG_B, ARG_C, ARG_D, INF_A, INF_B, INF_C, INF_D)) %>% 
  rowwise() %>% 
  mutate(value = mean(c_across(contains("_")))) %>% 
  ungroup() %>% 
  select(param, location, value, data)

spl2 <- read_csv(files[2], id = "data") %>% 
  pivot_wider(names_from = c(lang), values_from = value) %>% 
  unnest(cols = c(EN, ES)) %>% 
  rowwise() %>% 
  mutate(value = mean(c_across(c(EN, ES)))) %>% 
  ungroup() %>% 
  select(param, location, value, data)

lift <- read_csv(files[3], id = "data") %>% 
  pivot_wider(names_from = c(task), values_from = value) %>% 
  unnest(cols = c(`pre-test`, `post-test`)) %>% 
  rowwise() %>% 
  mutate(value = mean(c_across(contains("-test")))) %>% 
  ungroup() %>% 
  select(param, location, value, data)

cato <- read_csv(files[4], id = "data") %>% 
  pivot_wider(names_from = c(task, group), values_from = value) %>% 
  unnest() %>% 
  rowwise() %>% 
  mutate(value = mean(c_across(contains("_")))) %>% 
  ungroup() %>% 
  select(param, location, value, data)

c2l1 <- read_csv(files[5], id = "data") %>% 
  select(param, location, value, data)

ps <- bind_rows(plantra, spl2, lift, cato, c2l1) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(across(data, 
                str_remove_all, 
                "\\.|\\/|stanout|\\mog.csv"),
         theta = logit_scaled(prob)) %>% 
  select(-prob) %>%
  pivot_longer(beta:theta) %>%
  pivot_wider(names_from = c(data, location), 
              values_from = value, 
              names_sep = "_") %>% 
  unnest(contains("_"))

# Get main effects, interactions, nested effects
ps_effects <- ps %>%
  rowwise() %>%
  mutate(# Main effects
         DS_1 = mean(c_across(starts_with("lift_"))) - 
                mean(c_across(starts_with("spl2_"))),
         DS_2 = mean(c_across(starts_with("lift_"))) - 
                mean(c_across(starts_with("plantra_"))),
         DS_3 = mean(c_across(starts_with("spl2_"))) - 
                mean(c_across(starts_with("plantra_"))),
         DS_4 = mean(c_across(starts_with("lift_"))) - 
                mean(c_across(starts_with("cato_"))),
         DS_5 = mean(c_across(starts_with("plantra_"))) - 
                mean(c_across(starts_with("cato_"))),
         DS_6 = mean(c_across(starts_with("spl2_"))) - 
                mean(c_across(starts_with("cato_"))),
         DS_7 = mean(c_across(starts_with("c2l1_"))) - 
                mean(c_across(starts_with("plantra_"))),
         DS_8 = mean(c_across(starts_with("c2l1_"))) - 
                mean(c_across(starts_with("cato_"))),
         DS_9 = mean(c_across(starts_with("c2l1_"))) - 
                mean(c_across(starts_with("spl2_"))),
         DS_10 = mean(c_across(starts_with("c2l1_"))) - 
                mean(c_across(starts_with("lift_"))),

         TL_1 =  mean(c_across(contains(" sentence"))) - 
                mean(c_across(contains("_before word"))),
         TL_2 =  mean(c_across(contains("_before"))) - 
                mean(c_across(contains("_within word"))),
         # Two-way interactions
         `DS_1:TL_1` = (`lift_before sentence` - `lift_before word`) -
                       (`spl2_before sentence` - `spl2_before word`),

         `DS_2:TL_1` = (`lift_before sentence` - `lift_before word`) -
                       (`plantra_before sentence` - `plantra_before word`),

         `DS_3:TL_1` = (`spl2_before sentence` - `spl2_before word`) -
                       (`plantra_before sentence` - `plantra_before word`),

         `DS_4:TL_1` = (`lift_before sentence` - `lift_before word`) -
                       (`cato_before sentence` - `cato_before word`),

         `DS_5:TL_1` = (`plantra_before sentence` - `plantra_before word`) - 
                       (`cato_before sentence` - `cato_before word`),

         `DS_6:TL_1` = (`spl2_before sentence` - `spl2_before word`) -
                       (`cato_before sentence` - `cato_before word`),

         `DS_7:TL_1` = (`c2l1_before sentence` - `c2l1_before word`) -
                       (`plantra_before sentence` - `plantra_before word`),

         `DS_8:TL_1` = (`c2l1_before sentence` - `c2l1_before word`) -
                       (`cato_before sentence` - `cato_before word`),

         `DS_9:TL_1` = (`c2l1_before sentence` - `c2l1_before word`) -
                       (`spl2_before sentence` - `spl2_before word`),

         `DS_10:TL_1` = (`c2l1_before sentence` - `c2l1_before word`) -
                        (`lift_before sentence` - `lift_before word`),

         `DS_1:TL_2` = (mean(c_across(starts_with("lift_") & contains("_before"))) - 
                        mean(c_across(starts_with("spl2_") & contains("_before")))) -
                       (`lift_within word` - `spl2_within word`),

         `DS_2:TL_2` = (mean(c_across(starts_with("lift_") & contains("_before"))) - 
                        mean(c_across(starts_with("plantra_") & contains("_before")))) -
                       (`lift_within word` - `plantra_within word`),

         `DS_3:TL_2` = (mean(c_across(starts_with("spl2_") & contains("_before"))) - 
                        mean(c_across(starts_with("plantra_") & contains("_before")))) -
                       (`spl2_within word` - `plantra_within word`),

         `DS_4:TL_2` = (mean(c_across(starts_with("lift_") & contains("_before"))) - 
                        mean(c_across(starts_with("cato_") & contains("_before")))) -
                       (`lift_within word` - `cato_within word`),

         `DS_5:TL_2` = (mean(c_across(starts_with("plantra_") & contains("_before"))) - 
                        mean(c_across(starts_with("cato_") & contains("_before")))) -
                       (`plantra_within word` - `cato_within word`),

         `DS_6:TL_2` = (mean(c_across(starts_with("spl2_") & contains("_before"))) - 
                        mean(c_across(starts_with("cato_") & contains("_before")))) -
                       (`spl2_within word` - `cato_within word`),

         `DS_7:TL_2` = (mean(c_across(starts_with("c2l1_") & contains("_before"))) - 
                        mean(c_across(starts_with("plantra_") & contains("_before")))) -
                       (`c2l1_within word` - `plantra_within word`),
         `DS_8:TL_2` = (mean(c_across(starts_with("c2l1_") & contains("_before"))) - 
                        mean(c_across(starts_with("cato_") & contains("_before")))) -
                       (`c2l1_within word` - `cato_within word`),
         `DS_9:TL_2` = (mean(c_across(starts_with("c2l1_") & contains("_before"))) - 
                        mean(c_across(starts_with("spl2_") & contains("_before")))) -
                       (`c2l1_within word` - `spl2_within word`),
         `DS_10:TL_2` = (mean(c_across(starts_with("c2l1_") & contains("_before"))) - 
                        mean(c_across(starts_with("lift_") & contains("_before")))) -
                       (`c2l1_within word` - `lift_within word`)
         ) %>%
  ungroup() %>%
  mutate(
    # Locations effecs
    lift_sent = `lift_before sentence` - `lift_before word`,
    lift_word = `lift_before word` - `lift_within word`,
    spl2_sent = `spl2_before sentence` - `spl2_before word`,
    spl2_word = `spl2_before word` - `spl2_within word`,
    cato_sent = `cato_before sentence` - `cato_before word`,
    cato_word = `cato_before word` - `cato_within word`,
    c2l1_sent = `c2l1_before sentence` - `c2l1_before word`,
    c2l1_word = `c2l1_before word` - `c2l1_within word`,
    plantra_sent = `plantra_before sentence` - `plantra_before word`,
    plantra_word = `plantra_before word` - `plantra_within word`
    # data set effects
         # before_word_lift_spl2 = `lift_before word` - `spl2_before word`,
         # before_word_lift_plantra = `lift_before word` - `plantra_before word`,
         # before_word_spl2_plantra = `spl2_before word` - `plantra_before word`,
         # before_word_cato_spl2 = `cato_before word` - `spl2_before word`,
         # before_word_cato_plantra = `cato_before word` - `plantra_before word`,
         # before_word_cato_lift = `cato_before word` - `lift_before word`,
         # before_word_c2l1_plantra = `c2l1_before word` - `plantra_before word`,
         # before_word_c2l1_cato = `c2l1_before word` - `cato_before word`,
         # before_word_c2l1_spl2 = `c2l1_before word` - `spl2_before word`,
         # before_word_c2l1_lift = `c2l1_before word` - `lift_before word`,
         # 
         # before_sentence_lift_spl2 = `lift_before sentence` - `spl2_before sentence`,
         # before_sentence_lift_plantra = `lift_before sentence` - `plantra_before sentence`,
         # before_sentence_spl2_plantra = `spl2_before sentence` - `plantra_before sentence`,
         # before_sentence_cato_spl2 = `cato_before sentence` - `spl2_before sentence`,
         # before_sentence_cato_plantra = `cato_before sentence` - `plantra_before sentence`,
         # before_sentence_cato_lift = `cato_before sentence` - `lift_before sentence`,
         # before_sentence_c2l1_plantra = `c2l1_before sentence` - `plantra_before sentence`,
         # before_sentence_c2l1_cato = `c2l1_before sentence` - `cato_before sentence`,
         # before_sentence_c2l1_spl2 = `c2l1_before sentence` - `spl2_before sentence`,
         # before_sentence_c2l1_lift = `c2l1_before sentence` - `lift_before sentence`,
         # 
         # within_word_lift_spl2 = `lift_within word` - `spl2_within word`,
         # within_word_lift_plantra = `lift_within word` - `plantra_within word`,
         # within_word_spl2_plantra = `spl2_within word` - `plantra_within word`,
         # within_word_cato_spl2 = `cato_within word` - `spl2_within word`,
         # within_word_cato_plantra = `cato_within word` - `plantra_within word`,
         # within_word_cato_lift = `cato_within word` - `lift_within word`,
         # within_word_c2l1_plantra = `c2l1_within word` - `plantra_within word`,
         # within_word_c2l1_cato = `c2l1_within word` - `cato_within word`,
         # within_word_c2l1_spl2 = `c2l1_within word` - `spl2_within word`,
         # within_word_c2l1_lift = `c2l1_within word` - `lift_within word`
    ) %>%
  rename(parameter = name)

write_csv(ps_effects, "tables/ps_effects_acrossdata.csv")

#conda <- c(2, 3, 4, 1)
#condb <- c(4, 5, 6, 1)
#condc <- c(1, 2, 3, 4)
#condd <- c(4, 3, 2, 1)

#me1 <- (conda + condb)/2 - (condc + condd)/2
#me2 <- (conda + condc)/2 - (condb + condd)/2
#int <- (conda - condc) - (condb - condd)

```

```{r}
ps_effects <- read_csv("tables/ps_effects_acrossdata.csv")

# Main effects and interaction summary
ps_effects_summary <- ps_effects %>% 
#  mutate(across(-parameter, scale)) %>% 
  select(parameter,
         starts_with("DS"), 
         starts_with("TL")) %>%
  pivot_longer(-parameter) %>%
  group_by(parameter, name) %>%
  summarise(
    across(value, list(est = mean,
                       lower = lower,
                       upper = upper,
                       BF = BF), .names = "{.fn}")) %>%
  ungroup() %>%
  mutate(across(where(is.numeric), round, 2),
#         across(est, str_c, " [", lower, " -- ", upper, "]"),
         across(BF, ~ifelse( . > 100, "> 100", as.character(.))),
         across(name, recode_factor, 
                DS_1 = "Dataset 1 (LIFT, SPL2)",
                DS_2 = "Dataset 2 (LIFT, PLanTra)",
                DS_3 = "Dataset 3 (SPL2, PLanTra)",
                DS_4 = "Dataset 4 (CATO, SPL2)",
                DS_5 = "Dataset 5 (CATO, PLanTra)",
                DS_6 = "Dataset 6 (CATO, LIFT)",
                DS_7 = "Dataset 7 (C2L1, PLanTra)",
                DS_8 = "Dataset 8 (C2L1, CATO)",
                DS_9 = "Dataset 9 (C2L1, SPL2)",
                DS_10 = "Dataset 10 (C2L1, LIFT)",

                TL_1 = "Location 1 (before sentence, before word)",
                TL_2 = "Location 2 (before word / sentence, within word)",
               `DS_1:TL_1` = "Dataset 1 : Location 1",
               `DS_1:TL_2` = "Dataset 1 : Location 2",
               `DS_2:TL_1` = "Dataset 2 : Location 1",
               `DS_2:TL_2` = "Dataset 2 : Location 2",
               `DS_3:TL_1` = "Dataset 3 : Location 1",
               `DS_3:TL_2` = "Dataset 3 : Location 2",
               `DS_4:TL_1` = "Dataset 4 : Location 1",
               `DS_4:TL_2` = "Dataset 4 : Location 2",
               `DS_5:TL_1` = "Dataset 5 : Location 1",
               `DS_5:TL_2` = "Dataset 5 : Location 2",
               `DS_6:TL_1` = "Dataset 6 : Location 1",
               `DS_6:TL_2` = "Dataset 6 : Location 2",
               `DS_7:TL_1` = "Dataset 7 : Location 1",
               `DS_7:TL_2` = "Dataset 7 : Location 2",
               `DS_8:TL_1` = "Dataset 8 : Location 1",
               `DS_8:TL_2` = "Dataset 8 : Location 2",
               `DS_9:TL_1` = "Dataset 9 : Location 1",
               `DS_9:TL_2` = "Dataset 9 : Location 2",
               `DS_10:TL_1` = "Dataset 10 : Location 1",
               `DS_10:TL_2` = "Dataset 10 : Location 2",
               .ordered = T)) %>%
#  select(-lower,-upper) %>%
  pivot_wider(names_from = parameter, values_from = c(est,lower,upper,BF)) %>%
  arrange(name) %>% 
  select(Predictor = name, ends_with("beta"), ends_with("delta"), ends_with("theta"))

```

```{r effects, results='asis'}
ps_effects_summary %>%
  mutate(est_1 = "",
         est_2 = "",
         est_3 = "") %>%
  select(Predictor, est_1, BF_beta, est_2, BF_delta, est_3, BF_theta) %>%
  knitr::kable(align =c("l", rep("c", 6)), 
               col.names = c("Predictor", rep(c("Est.", "BF"), 3)),
               caption = "Mixture model results of transition duration with predictor estimates (main effects, interactions) for the distribution of short and hesitant transition durations (on log scale) and the probability of hesitant transitions (on logit scale). Estimates are shown with 95% PI. Dotted vertical line indicates 0.") %>%
  kable_styling(position = "center", full_width = T, font_size = 12) %>%
  column_spec(2, image = spec_pointrange(
    x = ps_effects_summary$est_beta, 
    xmin = ps_effects_summary$lower_beta, 
    xmax = ps_effects_summary$upper_beta, 
    vline = 0)
    ) %>% 
  column_spec(4, image = spec_pointrange(
    x = ps_effects_summary$est_delta, 
    xmin = ps_effects_summary$lower_delta, 
    xmax = ps_effects_summary$upper_delta, 
    vline = 0)
  ) %>% 
  column_spec(6, image = spec_pointrange(
    x = ps_effects_summary$est_theta, 
    xmin = ps_effects_summary$lower_theta, 
    xmax = ps_effects_summary$upper_theta, 
    vline = 0)
  ) %>% 
  footnote("Colon indicates interactions. PI is the probability interval. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", 
           threeparttable = T) %>%
  add_header_above(c("", 
                     "Short transition duration" = 2, 
                     "Slowdown for hesitant transitions" = 2, 
                     "Probability of hesitant transitions" = 2))  %>%
  group_rows("Main effects", 1, 12) %>%
  group_rows("Two-way interactions", 13, 32)  %>% 
  column_spec(1, width = "5.5cm") %>% 
  column_spec(2:7, width = "3cm")

```

```{r eval = F}
ps_effects <- read_csv("tables/ps_effects_acrossdata.csv")

cellmeans <- ps_effects %>% 
  select(parameter, contains(" ")) %>%
  pivot_longer(-parameter) %>%
  pivot_wider(names_from = parameter, values_from = value) %>%
  unnest() %>%
  mutate(delta = exp(beta + delta) - exp(beta),
         across(beta, exp),
         across(theta, ilogit)) %>%
  pivot_longer(cols = beta:theta, names_to = "param") %>% 
  group_by(name, param) %>%
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper),
                   .names = "{.fn}")) %>%
  ungroup() %>% 
  separate(name, into = c("data", "loc"), sep = "_") %>%
  mutate(across(data, recode, 
                "lift" = "LIFT", 
                "spl2" = "SPL2", 
                "plantra" = "PLanTra",
                "cato" = "CATO",
                "c2l1" = "C2L1")) %>% 
  pivot_wider(names_from = param, values_from = c(mean, lower, upper)) %>%
  mutate(across(c(ends_with("beta"), ends_with("delta")), round, 0),
         across(c(ends_with("beta"), ends_with("delta")), scales::comma, accuracy = 1),
         across(contains("theta"), dezero, 2),
         across(everything(), as.character)) %>%
  pivot_longer(contains("_"), 
               names_to = c(".value", "parameter"), 
               names_sep ="_") %>%
#  mutate(across(mean, str_c, " [", lower, " -- ", upper, "]")) %>%
#  select(-lower,-upper) %>% 
  pivot_wider(names_from = loc, values_from = c(mean, lower, upper))

tmp <- ps_effects %>% 
  select(parameter, ends_with("_sent"), ends_with("_word")) %>%
  pivot_longer(-parameter) %>%
  group_by(parameter, name) %>%
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper,
                        BF = BF),
                   .names = "{.fn}")) %>%
  ungroup() %>%
  mutate(across(where(is.numeric), round, 2),
#         across(mean, str_c, " [", lower, " -- ", upper, "]"),
         across(BF, ~ifelse(.>100, "> 100", as.character(.)))) %>%
  rename(diff = mean) %>% 
#  select(-lower,-upper) %>% 
  separate(name, into = c("data", "loc")) %>% 
  mutate(across(data, recode, 
                "lift" = "LIFT", 
                "spl2" = "SPL2", 
                "plantra" = "PLanTra",
                "cato" = "CATO",
                "c2l1" = "C2L1"),
         across(loc, recode, 
                "sent" = "before sentence - before word",
                "word" = "before word - within word")) %>% 
  right_join(cellmeans) %>% 
  arrange(parameter, data, loc) %>% 
  select(parameter, data, #`Transition location` = loc, 
         ends_with("before sentence"), 
         ends_with("before word"), 
         ends_with("within word"),
         Contrasts = loc, diff, lower, upper, BF)

write_csv(tmp, "tables/cellmeans_transdur.csv")
```

```{r cellmeans, results = 'asis'}
tmp <- read_csv("tables/cellmeans_transdur.csv")

cellmeans <- tmp %>% select(1:2, 
                            starts_with("mean_"), 
                            starts_with("lower_"),
                            starts_with("upper_")) %>% 
  pivot_longer(-1:-2) %>% 
  unique() %>% 
  arrange(parameter, data) %>% 
  select(-parameter, -data)

table <- tmp %>% select(-contains("_")) %>% 
  arrange(parameter, Contrasts, lower) %>% 
  pivot_wider(names_from = parameter, 
              values_from = c(diff, lower, upper, BF))

table %>% 
  mutate(est_1 = "",
         est_2 = "",
         est_3 = "") %>%  
  select(data, #Contrasts, 
         est_1, BF_beta,
         est_2, BF_delta,
         est_3, BF_theta) %>% 
  knitr::kable(
  align =c("l", rep("c", 6)),
  col.names = c("Data set", rep(c("Est.", "BF"),3)),
  caption = "By-data set differences between transition locations for transition duration estimates inferred from mixture model. Differences are shown on the log scale (for durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets. Dotted line indicates 0.") %>%
#  collapse_rows(columns = 1, valign = "middle") %>% 
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 9) %>%
    column_spec(2, image = spec_pointrange(
    x = table$diff_beta, 
    xmin = table$lower_beta, 
    xmax = table$upper_beta, 
    vline = 0)
    ) %>% 
  column_spec(4, image = spec_pointrange(
    x = table$diff_delta, 
    xmin = table$lower_delta, 
    xmax = table$upper_delta, 
    vline = 0)
  ) %>% 
  column_spec(6, image = spec_pointrange(
    x = table$diff_theta, 
    xmin = table$lower_theta, 
    xmax = table$upper_theta, 
    vline = 0)
  ) %>% 
  footnote("PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", 
           threeparttable = T) %>% 
  add_header_above(c(" " = 1, "Short interval durations" = 2, "Slowdown for hesitant transitions" = 2, "Probability of hesitant transitions" = 2)) %>% 
#  group_rows("Short transition duration", 1, 9) %>%
#  group_rows("Slowdown for hesitant transitions", 10, 18) %>%
#  group_rows("Probability of hesitant transitions", 19, 27) #%>%
#  column_spec(1, "1.75cm") %>%
#  column_spec(2, "4cm") %>% 
  group_rows("Contrast: before sentence -- before word", 1, 5) %>%
  group_rows("Contrast: before word -- within word", 6, 10)

#  column_spec(2:4, "1.75cm") %>% 
#  column_spec(5:6, "1.75cm") %>% 
#  column_spec(7, "1cm") 
```

# References