### 2023-07-24 report/appendix_a.Rmd

## Fit to data

Model fit was assessed by testing to what extent simulations obtained from the model fit the observed data. From each model we obtained 100 simulations (predictions). These can be found in Figure \ref{fig:prediction}. The model predictions fitted the data substantially better for the mixture models. The worse fit was observed for the unimodal Gaussian model which predicted large amounts of negative keystroke intervals which are implausible. Using a log-normal distribution removed these negative keystroke intervals.

```{r prediction, fig.cap="Fit to data. Comparison of 100 simulated (predicted) sets of data to observed data summarised by model. For illustration the x-axis was truncated at 2 secs.", eval = T, fig.height=8, fig.width=6.5}

run <- F

if(run){
  theme_set(theme_few(base_size = 9) +
              theme(strip.background = element_blank(),
                    legend.position = "top",
                    legend.justification = "right",
                    panel.grid = element_blank(),
                    panel.background = element_rect(fill = "transparent"), # bg of the panel
                    plot.background = element_rect(fill = "transparent", color = NA),
                    axis.text.y = element_blank(),
                    axis.ticks.y = element_blank(),
                    axis.title = element_blank()))
  file <- list.files("../stanout", pattern = "sims", full.names = T, recursive = T)
  ds <- str_remove_all(file, "\\.|\\/|stanout|all_sims.csv") %>% 
    str_to_upper() %>% 
    str_replace(., "PLANTRA", "PLanTra")
  ds <- ds[-6]
  plots <- list()
  plots <- map(1:length(ds), ~get_pdens_plot(file[.], max_iki = 2000, ds = ds[.])) 
  plots_fin <- plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] / plots[[6]] + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

  plots_fin <- grid.arrange(patchworkGrob(plots_fin), 
                            left = "Density",
                            bottom = "Transition duration in secs")

  ggsave("figures/fitplots.pdf", plots_fin, height = 8.5, width = 6.25)
}

knitr::include_graphics("figures/fitplots.pdf")
```