### 2023-07-24 report/appendix_b.Rmd

## Posterior parameter estimates

```{r}
# get posterior
file <- list.files(path = "../stanout/cato", 
                   pattern = "mogbetaun.+.csv$",
                   full.names = T)

ps_cato <- read_csv(file) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(task, group, location, name)) %>% 
  mutate(group = str_c("Task: ", task, "; Group: ", group),
         ds = "CATO") %>% 
  select(-task)

file <- list.files(path = "../stanout/c2l1", 
                   pattern = "mogbetaun.*.csv$",
                   full.names = T)

ps_c2l1 <- read_csv(file) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(location, name)) %>% 
  mutate(ds = "C2L1",
         group = "1")

file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetaun.*.csv$",
                   full.names = T)

ps_gunn <- read_csv(file) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(location, xn, name)) %>% 
  mutate(ds = "GUNNEXP2") %>% 
  rename(group = xn)

file <- list.files(path = "../stanout/lift", 
                   pattern = "mogbetaun.+.csv$",
                   full.names = T)

ps_lift <- read_csv(file) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(genre, topic, location, name)) %>% 
  mutate(group = str_c("Topic: ", topic, "; Genre: ", genre),
         ds = "LIFT") %>% 
  select(-topic, -genre)

file <- list.files(path = "../stanout/plantra", 
                   pattern = "mogbetaun.+.csv$",
                   full.names = T)

ps_plantra <- read_csv(file) %>% 
  mutate(across(task, factor, 
                levels = c("pre-test", "post-test"), 
                ordered = T)) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(task, location, name)) %>% 
  mutate(ds = "PLanTra") %>% 
  rename(group = task)

file1 <- list.files(path = "../stanout/spl2", 
                    pattern = "mogbetaun.+.csv$",
                    full.names = T)

file2 <- list.files(path = "../stanout/spl2_shift", 
                    pattern = "mogbetaun.+.csv$",
                    full.names = T)

ps_spl2 <- map_dfr(c(file1, file2), ~read_csv(.) %>% 
                     mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode(., spl2 = "SPL2 (_^[shift])",
                              spl2_shift = "SPL2 (_^[shift] + C)"))) %>% 
  filter(param %in% c("beta", "delta", "prob")) %>% 
  mutate(across(lang, recode, "EN" = "L1", "ES" = "L2")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, delta, prob)) %>% 
  mutate(beta2 = beta + delta,
         across(c(beta2, beta), ~exp(.)),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper),
                   .names = "{.fn}"),
            .by = c(lang, location, name, data)) %>% 
  mutate(ds = "SPL2",
         group = str_c("Data: ", data, "; Language: ", lang)) %>% 
  select(-lang)

# Combine all
ps <- bind_rows(ps_cato, ps_c2l1, ps_gunn, ps_lift, ps_plantra, ps_spl2)
```

```{r eval = F}
theme_set(theme_few(base_size = 11) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

pdodge <- position_dodge(-.5)
beta_plots <- list()
delta_plots <- list()
theta_plots <- list()
datasets <- unique(ps$ds)
size <- 2
width <- 0
linewidth <- .35

beta_plots <- map(datasets, ~filter(ps, 
                                    name == "beta", 
                                    ds == .x) %>% 
                    mutate(across(c(ds,group), ~factor(.)),
                           name = beta_label) %>% 
                    ggplot(aes(y = mean, 
                               ymin = lower, 
                               ymax = upper,
                               x = location,
                               shape = group)) +
                    geom_errorbar(width = width,
                                  linewidth = linewidth,
                                  alpha = .75, 
                                  position = pdodge) +
                    geom_point(aes(colour = group), 
                               size = size, 
                               position = pdodge) +
                    facet_grid(~name) +
                    scale_shape_manual(values = 1:length(filter(ps, ds == .x) %>% 
                                                           pull(group) %>% 
                                                           factor())) +
                    scale_colour_colorblind() +
                    scale_y_continuous(labels = scales::comma) +
                    labs(subtitle = .x) +
                    theme(axis.title = element_blank(),
                          legend.title = element_blank()) +
                    guides(colour=guide_legend(nrow=2,byrow=TRUE),
                           shape = guide_legend(nrow=2,byrow=TRUE)))

delta_plots <- map(datasets, ~filter(ps, 
                                     name == "delta", 
                                     ds == .x) %>% 
                     mutate(across(c(ds,group), ~factor(.)),
                            across(c(mean, lower, upper), ~ . /1000),
                            name = delta_label) %>% 
                     ggplot(aes(y = mean, 
                                ymin = lower, 
                                ymax = upper,
                                x = location,
                                shape = group)) +
                     geom_errorbar(width = width, 
                                   alpha = .75, 
                                   linewidth = linewidth,
                                   position = pdodge) +
                     geom_point(aes(colour = group), 
                                size = size, 
                                position = pdodge) +
                     facet_grid(~name) +
                     scale_shape_manual(values = 1:length(filter(ps, ds == .x) %>% 
                                                            pull(group) %>% 
                                                            factor())) +
                     scale_colour_colorblind() +
                     scale_y_log10(labels = scales::comma) +
                     theme(axis.title = element_blank(),
                           legend.title = element_blank()) +
                     guides(colour=guide_legend(nrow=2,byrow=TRUE),
                            shape = guide_legend(nrow=2,byrow=TRUE)))

theta_plots <- map(datasets, ~filter(ps, 
                                     name == "prob", 
                                     ds == .x) %>% 
                     mutate(across(c(ds,group), ~factor(.)),
                            name = theta_label) %>% 
                     ggplot(aes(y = mean, 
                                ymin = lower, 
                                ymax = upper,
                                x = location,
                                shape = group)) +
                     geom_errorbar(width = width,
                                   linewidth = linewidth,
                                   alpha = .75, 
                                   position = pdodge) +
                     geom_point(aes(colour = group), 
                                size = size, 
                                position = pdodge) +
                     facet_grid(~name) +
                     scale_shape_manual(values = 1:length(filter(ps, ds == .x) %>% 
                                                            pull(group) %>% 
                                                            factor())) +
                     scale_colour_colorblind() +
                     scale_y_continuous(labels = dezero_plot, 
                                        limits = c(0, 1)) +
                     theme(axis.title = element_blank(),
                           legend.title = element_blank()) +
                     guides(colour=guide_legend(nrow=2,byrow=TRUE),
                            shape = guide_legend(nrow=2,byrow=TRUE)))

# Combine parameter plots 
plot_param <- list()
plot_param <- map(1:6, ~beta_plots[[.]] + delta_plots[[.]] + theta_plots[[.]] +
                    plot_layout(guides = "collect") &
                    theme(legend.position = "bottom"))

plot_param[[2]] <- plot_param[[2]] & theme(legend.position = "none")
plot_param[[3]] <- plot_param[[3]] & guides(colour=guide_legend(nrow=1,byrow=TRUE),
                                            shape = guide_legend(nrow=1,byrow=TRUE))

plot_param[[5]] <- plot_param[[5]] & guides(colour=guide_legend(nrow=1,
                                                                byrow=TRUE,
                                                                reverse=TRUE),
                                            shape = guide_legend(nrow=1,
                                                                 byrow=TRUE,
                                                                 reverse=TRUE))

plot_fin_1 <- plot_param[[1]] / plot_param[[2]] / plot_param[[3]]
plot_fin_1 <- grid.arrange(patchworkGrob(plot_fin_1), 
                           left = "Posterior estimate with 95% PIs",
                           bottom = "Transition location")
ggsave("figures/psplots1.pdf", plot_fin_1, height = 8, width = 7)

plot_fin_2 <- plot_param[[4]] / plot_param[[5]] / plot_param[[6]]
plot_fin_2 <- grid.arrange(patchworkGrob(plot_fin_2), 
                           left = "Posterior estimate with 95% PIs",
                           bottom = "Transition location")
ggsave("figures/psplots2.pdf", plot_fin_2, height = 8, width = 7)
```

\begin{figure}[!htb]
\centering
\includegraphics{figures/psplots1.pdf}
\caption{Posterior parameter distribution}
\end{figure}
\newpage
\begin{figure}[!htb]
\ContinuedFloat
\captionsetup{list=off,format=cont}
\centering
\includegraphics{figures/psplots2.pdf}
\label{fig:fullps1}
\caption{Posterior parameter distribution}
\end{figure}