### 2023-07-24 report/appendix_g.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\hypertarget{pre-post-test-plantra}{%
\subsection{Pre-post test (PLanTra)}\label{pre-post-test-plantra}}

The pre-post test effect for the PLanTra dataset is reported in Table
\ref{tab:retesteffect}. Transition durations were shorter at before-word
locations for the post-test. Evidence was negligible for any other
comparisons.

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.}
\end{TableNotes}

\footnotesize{

\begin{longtable}{lrrrr}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:retesteffect}Mixture model estimates for post-test effect. Cell means are shown for the pre-test and post-test in msecs for fluent key-transitions, the slowdown for long transitions and the probability of hesitant transitions. The effect for post-test is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95\% PIs in brackets.}\\
\toprule
Transition location & \multicolumn{1}{c}{Pre-test} & \multicolumn{1}{c}{Post-test} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:retesteffect} continued}}\\
\toprule
Transition location & \multicolumn{1}{c}{Pre-test} & \multicolumn{1}{c}{Post-test} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endhead
Fluent transitions &  &  &  & \\
\ \ \ before sentence & 170 [156, 185] & 175 [161, 189] & -0.03 [-0.11, 0.05] & 0.05\\
\ \ \ before word & 176 [166, 186] & 166 [156, 176] & 0.06 [0.03, 0.09] & 21.33\\
\ \ \ within word & 145 [137, 154] & 146 [138, 154] & 0 [-0.03, 0.02] & 0.01\\
Hesitation duration &  &  &  & \\
\ \ \ before sentence & 1,029 [827, 1,260] & 1,107 [877, 1,364] & -0.04 [-0.25, 0.17] & 0.12\\
\ \ \ before word & 480 [426, 540] & 463 [410, 520] & -0.01 [-0.12, 0.09] & 0.05\\
\ \ \ within word & 224 [174, 284] & 242 [188, 306] & -0.05 [-0.25, 0.16] & 0.11\\
Hesitation probability &  &  &  & \\
\ \ \ before sentence & .52 [.44, .60] & .48 [.40, .55] & 0.17 [-0.23, 0.58] & 0.29\\
\ \ \ before word & .48 [.43, .54] & .45 [.39, .50] & 0.15 [-0.16, 0.45] & 0.24\\
\ \ \ within word & .13 [.10, .16] & .12 [.09, .14] & 0.13 [-0.24, 0.49] & 0.24\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

}

\end{ThreePartTable}
\end{center}
\end{appendix}