### 2023-12-09 report/appendix_genre_constr.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\hypertarget{genre-effect-lift}{%
\subsection{Genre effect (LIFT)}\label{genre-effect-lift}}

From the LIFT data, we compared the difference between genres, i.e.~when
writing an informative text as opposed to writing an argumentative text.
The results are shown in Table \ref{tab:genreeffect}. Cellmeans and
differences were average across writing topic. Evidence for a genre
comparisons was negligible.

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.}
\end{TableNotes}

\footnotesize{

\begin{longtable}{lrrrr}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:genreeffect}Mixture model estimates for genre effect. Cell means are shown for argumentative and informative texts in msecs for the hesitation slowdown and the probability of hesitant transitions. The effect for genre is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95\% PIs in brackets.}\\
\toprule
Transition location & \multicolumn{1}{c}{Argumentative} & \multicolumn{1}{c}{Informative} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:genreeffect} continued}}\\
\toprule
Transition location & \multicolumn{1}{c}{Argumentative} & \multicolumn{1}{c}{Informative} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endhead
Hesitation duration &  &  &  & \\
\ \ \ before sentence & 373 [301, 454] & 411 [330, 504] & -0.07 [-0.33, 0.19] & 0.16\\
\ \ \ before word & 287 [250, 326] & 284 [248, 323] & 0.01 [-0.15, 0.16] & 0.09\\
\ \ \ within word & 199 [118, 306] & 246 [155, 361] & -0.12 [-0.6, 0.36] & 0.29\\
Hesitation probability &  &  &  & \\
\ \ \ before sentence & .35 [.28, .43] & .34 [.27, .42] & 0.05 [-0.57, 0.73] & 0.36\\
\ \ \ before word & .46 [.39, .54] & .48 [.41, .55] & -0.05 [-0.51, 0.44] & 0.25\\
\ \ \ within word & .07 [.05, .10] & .07 [.05, .10] & -0.02 [-0.81, 0.76] & 0.46\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

}

\end{ThreePartTable}
\end{center}
\end{appendix}