### 2023-12-09 report/appendix_l2_constr.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\hypertarget{l2-effect-spl2}{%
\subsection{L2 effect (SPL2)}\label{l2-effect-spl2}}

For the SPL2 data (only the \texttt{\_\^{}{[}shift{]}}
sentence-transitions) we calculated the L2 effect (i.e.~the difference
between writing in L2 and L1). The results can be found in Table
\ref{tab:l2effect}. The results show longer hesitation duration and more
pauses across all transition locations when writing in L2. Only
within-word transitions did show negligible evidence for an increased
hesitation duration in L2.

\begin{table}[b]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:l2effect}Mixture model estimates for language effect. Cell means are shown the slowdown for hesitant transitions when writing in L1 and L2 and the probability of hesitant transitions. The language difference is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95\% PIs in brackets.}

\small{

\begin{tabular}{lrrrr}
\toprule
Transition location & \multicolumn{1}{c}{L1} & \multicolumn{1}{c}{L2} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
Hesitation duration &  &  &  & \\
\ \ \ before sentence & 1,281 [1,082, 1,496] & 1,802 [1,557, 2,073] & 0.31 [0.17, 0.45] & > 100\\
\ \ \ before word & 289 [254, 327] & 560 [507, 617] & 0.47 [0.39, 0.55] & > 100\\
\ \ \ within word & 120 [53, 222] & 202 [156, 258] & 0.27 [-0.07, 0.55] & 0.68\\
Hesitation probability &  &  &  & \\
\ \ \ before sentence & .78 [.69, .86] & .92 [.87, .96] & 1.23 [0.52, 1.97] & > 100\\
\ \ \ before word & .51 [.41, .62] & .89 [.83, .93] & 2.02 [1.4, 2.66] & > 100\\
\ \ \ within word & .04 [.02, .07] & .15 [.10, .22] & 1.52 [0.78, 2.29] & > 100\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}
\end{appendix}