### 2023-12-09 report/appendix_masking_constr.Rmd

## Masking effect (CATO, GUNNEXP2)

Studies associated with two datasets (CATO, GUNNEXP2) investigated to what extent masking the previously written text affects keystroke behaviour. The mixture model results for the effect of masking is shown in Table \ref{tab:maskingeffect}. There was some evidence for longer before-sentence hesitation durations when the text was masked in non-dyslexic writers. Evidence for all other comparisons was negligible.

\blandscape
```{r maskingeffect, results = 'asis'}
ps_cato_2 <- ps_cato %>% 
  separate(group, into = c("mask","group"), sep = "; ") %>% 
  mutate(across(c(mask, group), ~str_remove(., "Task: |Group: ")))

cellmeans_mask <- ps_gunn %>% 
  rename(mask = group) %>%
  mutate(group = "non dyslexic") %>% 
  bind_rows(ps_cato_2) %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, group, mask, ds), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = mask, 
              values_from = est)

file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetaco.*.csv$",
                   full.names = T)

d_gunn <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value)),
         ds = "gunnexp2",
         across(param, ~recode(., "prob" = "theta")),
         group = "non dyslexic") %>% 
  rename(mask = xn)

# get posterior
file <- list.files(path = "../stanout/cato", 
                   pattern = "mogbetaco.+.csv$",
                   full.names = T)

d_cato <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "prob")) %>% 
  mutate(ds = "cato") %>% 
  rename(mask = task)

tmp <- bind_rows(d_gunn, d_cato) %>% 
  pivot_wider(names_from = mask, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param, group, ds)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob")),
         across(ds, ~str_to_upper(.))) %>% 
  left_join(cellmeans_mask, by = join_by(location, param, group, ds)) %>% 
  arrange(param, location, ds, group) %>% 
  select(param, `Transition location` = location,
         `Dataset` = ds, Group = group, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "tables/masking_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "l", "l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for masking effect. Cell means are shown for the masked and unmasked writing task in msecs for the slowdown for long transitions and the probability of hesitant transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Hesitation duration" = 1:9,
            "Hesitattion probability" = 10:18) )

```
\elandscape