### 2023-12-09 report/appendix_masking_constr.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\hypertarget{masking-effect-cato-gunnexp2}{%
\subsection{Masking effect (CATO,
GUNNEXP2)}\label{masking-effect-cato-gunnexp2}}

Studies associated with two datasets (CATO, GUNNEXP2) investigated to
what extent masking the previously written text affects keystroke
behaviour. The mixture model results for the effect of masking is shown
in Table \ref{tab:maskingeffect}. There was some evidence for longer
before-sentence hesitation durations when the text was masked in
non-dyslexic writers. Evidence for all other comparisons was negligible.

\blandscape

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.}
\end{TableNotes}

\footnotesize{

\begin{longtable}{lllrrrr}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:maskingeffect}Mixture model estimates for masking effect. Cell means are shown for the masked and unmasked writing task in msecs for the slowdown for long transitions and the probability of hesitant transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95\% PIs in brackets.}\\
\toprule
Transition location & \multicolumn{1}{c}{Dataset} & \multicolumn{1}{c}{Group} & \multicolumn{1}{c}{Unmasked} & \multicolumn{1}{c}{Masked} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:maskingeffect} continued}}\\
\toprule
Transition location & \multicolumn{1}{c}{Dataset} & \multicolumn{1}{c}{Group} & \multicolumn{1}{c}{Unmasked} & \multicolumn{1}{c}{Masked} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endhead
Hesitation duration &  &  &  &  &  & \\
\ \ \ before sentence & CATO & dyslexic & 2,231 [1,543, 3,111] & 1,680 [1,239, 2,205] & -0.25 [-0.64, 0.13] & 0.44\\
\ \ \ before sentence & CATO & non dyslexic & 1,733 [1,241, 2,325] & 1,502 [1,134, 1,942] & -0.12 [-0.47, 0.22] & 0.23\\
\ \ \ before sentence & GUNNEXP2 & non dyslexic & 1,124 [954, 1,307] & 1,418 [1,218, 1,634] & 0.21 [0.05, 0.36] & 2.59\\
\ \ \ before word & CATO & dyslexic & 444 [405, 485] & 407 [369, 447] & -0.06 [-0.13, 0.01] & 0.14\\
\ \ \ before word & CATO & non dyslexic & 345 [312, 381] & 326 [295, 361] & -0.04 [-0.11, 0.04] & 0.06\\
\ \ \ before word & GUNNEXP2 & non dyslexic & 363 [325, 403] & 379 [341, 420] & 0.03 [-0.05, 0.11] & 0.05\\
\ \ \ within word & CATO & dyslexic & 278 [166, 400] & 177 [116, 252] & -0.24 [-0.52, 0.06] & 0.57\\
\ \ \ within word & CATO & non dyslexic & 409 [244, 593] & 392 [230, 591] & -0.03 [-0.42, 0.38] & 0.2\\
\ \ \ within word & GUNNEXP2 & non dyslexic & 182 [105, 279] & 209 [135, 306] & 0.08 [-0.24, 0.4] & 0.19\\
Hesitattion probability &  &  &  &  &  & \\
\ \ \ before sentence & CATO & dyslexic & .60 [.38, .79] & .80 [.60, .93] & 0.2 [-0.06, 0.46] & 0.45\\
\ \ \ before sentence & CATO & non dyslexic & .55 [.34, .76] & .68 [.47, .84] & 0.13 [-0.16, 0.4] & 0.21\\
\ \ \ before sentence & GUNNEXP2 & non dyslexic & .88 [.82, .93] & .90 [.85, .94] & 0.16 [-0.46, 0.79] & 0.36\\
\ \ \ before word & CATO & dyslexic & .75 [.64, .85] & .72 [.59, .82] & -0.04 [-0.19, 0.12] & 0.09\\
\ \ \ before word & CATO & non dyslexic & .64 [.51, .75] & .62 [.49, .74] & -0.02 [-0.19, 0.16] & 0.09\\
\ \ \ before word & GUNNEXP2 & non dyslexic & .47 [.40, .54] & .45 [.38, .52] & -0.06 [-0.46, 0.35] & 0.21\\
\ \ \ within word & CATO & dyslexic & .07 [.03, .12] & .11 [.06, .19] & 0.05 [-0.02, 0.13] & 0.08\\
\ \ \ within word & CATO & non dyslexic & .03 [.02, .06] & .03 [.02, .06] & 0 [-0.03, 0.03] & 0.01\\
\ \ \ within word & GUNNEXP2 & non dyslexic & .05 [.03, .08] & .04 [.03, .06] & -0.29 [-0.88, 0.28] & 0.48\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

}

\end{ThreePartTable}
\end{center}
\elandscape
\end{appendix}