### 2023-12-09 report/appendix_prepost_constr.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\hypertarget{pre-post-test-plantra}{%
\subsection{Pre-post test (PLanTra)}\label{pre-post-test-plantra}}

The pre-post test effect for the PLanTra dataset is reported in Table
\ref{tab:retesteffect}. Evidence for all comparisons was negligible.

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.}
\end{TableNotes}

\footnotesize{

\begin{longtable}{lrrrr}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:retesteffect}Mixture model estimates for post-test effect. Cell means are shown for the pre-test and post-test in msecs for the slowdown for hesitant transitions and the probability of hesitant transitions. The effect for post-test is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95\% PIs in brackets.}\\
\toprule
Transition location & \multicolumn{1}{c}{Pre-test} & \multicolumn{1}{c}{Post-test} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:retesteffect} continued}}\\
\toprule
Transition location & \multicolumn{1}{c}{Pre-test} & \multicolumn{1}{c}{Post-test} & \multicolumn{1}{c}{Difference} & \multicolumn{1}{c}{BF}\\
\midrule
\endhead
Hesitation duration &  &  &  & \\
\ \ \ before sentence & 879 [723, 1,052] & 909 [740, 1,100] & -0.03 [-0.23, 0.17] & 0.1\\
\ \ \ before word & 438 [395, 485] & 434 [389, 483] & 0.01 [-0.08, 0.1] & 0.05\\
\ \ \ within word & 245 [180, 323] & 269 [201, 353] & -0.06 [-0.3, 0.18] & 0.14\\
Hesitation probability &  &  &  & \\
\ \ \ before sentence & .58 [.50, .66] & .54 [.47, .62] & 0.15 [-0.26, 0.58] & 0.27\\
\ \ \ before word & .56 [.50, .61] & .49 [.43, .55] & 0.27 [-0.04, 0.59] & 0.72\\
\ \ \ within word & .11 [.08, .14] & .09 [.07, .12] & 0.13 [-0.27, 0.54] & 0.26\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

}

\end{ThreePartTable}
\end{center}
\end{appendix}