### 2023-12-09 report/appendix_shift_constr.Rmd

## Key-combination effect (constrained mixture model)

The analysed datasets differ to the extent that keystroke intervals at before sentence location sometimes did (PLanTra, LIFT) or did not (CATO, C2L1, SPL2, GUNNEXP2) scope over the character following the shift key. In other words, the pause before sentences summed across two key intervals in the PLanTra and LIFT data, namely `_^[shift]^C` but only involved one key interval, namely `_^[shift]` for the remaining datasets. Therefore, longer or more frequent pauses at before-sentence locations compared to before-word locations can be explained without reference to linguistic edges.

<!-- Also there is a possibility that some inconsistencies in our findings can be explained on the basis of including the keystroke following shift.  -->

Therefore we compared whether the different patterns can be explain on the basis of the additional keystroke involved in before-sentence transitions. We compared the SPL2 data including and excluding the keystroke after shift. Although we modelled all transition locations, we present only before-sentence transitions below as there was, as one would expect, no difference at word locations. The results of this comparison can be found in Table \ref{tab:shiftcellmeans}.

```{r shiftcellmeans, results = 'asis'}
# cell means and data set difference
cellmeans <- ps_spl2_both %>% 
  separate(group, into = c("group", "lang"), sep = "; ") %>%
  select(-ds, -group) %>%
  mutate(across(lang, ~str_remove(., "Language: ")),
         across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(lang, location, data), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = data, 
              values_from = est) %>% 
  mutate(across(c(location, lang), ~ifelse(param == "beta", "overall", .))) %>% 
  unique()

file1 <- list.files(path = "../stanout/spl2", 
                    pattern = "mogbetacon.+.csv$",
                    full.names = T)

file2 <- list.files(path = "../stanout/spl2_shift", 
                    pattern = "mogbetacon.+.csv$",
                    full.names = T)

tmp <- map_dfr(c(file1, file2), ~read_csv(.x) %>% 
                 mutate(data = .x)) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  mutate(across(lang, recode, "EN" = "L1", "ES" = "L2"),
         across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .))) %>% 
  pivot_wider(names_from = data, values_from = value) %>% 
  unnest(c(spl2, spl2_shift)) %>% 
  mutate(diff = spl2_shift - spl2) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(lang, location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, recode, theta = "prob")) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location, lang) %>% 
  filter(location %in% c("overall", "before sentence")) %>% 
  select(param, 
         Language = lang,
         `_^[shift] + C` = ends_with(" + C)"),
         `_^[shift]` = ends_with("FT])"),
         "Difference" = mean, BF)

write_csv(tmp, "tables/shift_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for key transitions immediately preceding a sentence. Cell means are shown for transitions that do and do not involve the transition to the character following shift in msecs for the slowdown for long transitions and the probability of hesitant transitions. The difference for including the transition duration to the character after shift is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          font_size = "footnotesize",
          longtable = T,
          stub_indents = list("Fluent transitions" = 1,
                              "Hesitation duration" = 2:3,
                              "Hesitation probability" = 4:5) )

```

The fluent transition duration and hesitation duration were not affected by whether or not the sentence-initial transition include the character following shift for neither language. However, we found strong evidence for an increased hesitation probability -- in both -- languages when the before-sentence key transition included the character following the shift key. Notably, the hesitation probability increased ceiling. It is not entirely surprised that the constrained mixture model identifies essentially all before-sentence transition as pauses: the constrained model does not distinguish between transition locations for the duration of fluent transitions. Therefore essentially all before-sentence transitions are identified as hesitation as they always include two keystrokes while all other transition location (which is the majority of the data) do not.