### 2024-04-27 report/appendix/postparam_constr.Rmd

# Posterior parameter estimates

```{r eval = F}

theme_set(theme_few(base_size = 11) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

pdodge <- position_dodge(-.5)
beta_plots <- list()
delta_plots <- list()
theta_plots <- list()
datasets <- unique(ps$ds)
size <- 2
width <- 0
linewidth <- .35
beta_label <- "Fluent transitions\n(in msecs)"

beta_plots <- map(datasets, ~filter(ps, 
                                    name == "beta", 
                                    location == location[1],
                                    ds == .x) %>% 
                    mutate(name = beta_label,
                           group == "overall",
                           location = "overall") %>% 
                    unique() %>% 
                    ggplot(aes(y = mean, 
                               ymin = lower, 
                               ymax = upper,
                               x = location)) +
                    geom_errorbar(width = width,
                                  linewidth = linewidth,
                                  alpha = .75,
                                  position = pdodge) +
                    geom_point(size = size,
                               position = pdodge) +
                    facet_grid(~name) +
                    scale_colour_colorblind() +
                    scale_y_continuous(labels = scales::comma) +
                    labs(subtitle = .x) +
                    theme(axis.title = element_blank(),
                          legend.title = element_blank()) +
                    guides(colour=guide_legend(nrow=1,byrow=TRUE),
                           shape = guide_legend(nrow=1,byrow=TRUE)))

delta_plots <- map(datasets, ~filter(ps, 
                                     name == "delta", 
                                     ds == .x) %>% 
                     mutate(across(c(ds,group), ~factor(.)),
                            across(c(mean, lower, upper), ~ . /1000),
                            name = delta_label) %>% 
                     ggplot(aes(y = mean, 
                                ymin = lower, 
                                ymax = upper,
                                x = location,
                                shape = group)) +
                     geom_errorbar(width = width, 
                                   alpha = .75, 
                                   linewidth = linewidth,
                                   position = pdodge) +
                     geom_point(aes(colour = group), 
                                size = size, 
                                position = pdodge) +
                     facet_grid(~name) +
                     scale_shape_manual(values = 1:length(filter(ps, ds == .x) %>% 
                                                            pull(group) %>% 
                                                            factor())) +
                     scale_colour_colorblind() +
                     scale_y_log10(labels = scales::comma) +
                     theme(axis.title = element_blank(),
                           legend.title = element_blank()) +
                     guides(colour=guide_legend(nrow=2,byrow=TRUE),
                            shape = guide_legend(nrow=2,byrow=TRUE)))

theta_plots <- map(datasets, ~filter(ps, 
                                     name == "prob", 
                                     ds == .x) %>% 
                     mutate(across(c(ds,group), ~factor(.)),
                            name = theta_label) %>% 
                     ggplot(aes(y = mean, 
                                ymin = lower, 
                                ymax = upper,
                                x = location,
                                shape = group)) +
                     geom_errorbar(width = width,
                                   linewidth = linewidth,
                                   alpha = .75, 
                                   position = pdodge) +
                     geom_point(aes(colour = group), 
                                size = size, 
                                position = pdodge) +
                     facet_grid(~name) +
                     scale_shape_manual(values = 1:length(filter(ps, ds == .x) %>% 
                                                            pull(group) %>% 
                                                            factor())) +
                     scale_colour_colorblind() +
                     scale_y_continuous(labels = dezero_plot, 
                                        limits = c(0, 1)) +
                     theme(axis.title = element_blank(),
                           legend.title = element_blank()) +
                     guides(colour=guide_legend(nrow=2,byrow=TRUE),
                            shape = guide_legend(nrow=2,byrow=TRUE)))

# Combine parameter plots 
plot_param <- list()
plot_param <- map(1:6, ~beta_plots[[.]] + delta_plots[[.]] + theta_plots[[.]] +
                    plot_layout(guides = "collect", widths = c(.45, 1, 1)) &
                    theme(legend.position = "bottom"))

plot_param[[2]] <- plot_param[[2]] & theme(legend.position = "none")
plot_param[[3]] <- plot_param[[3]] & guides(colour=guide_legend(nrow=1,byrow=TRUE),
                                            shape = guide_legend(nrow=1,byrow=TRUE))

plot_param[[5]] <- plot_param[[5]] & guides(colour=guide_legend(nrow=1,
                                                                byrow=TRUE,
                                                                reverse=TRUE),
                                            shape = guide_legend(nrow=1,
                                                                 byrow=TRUE,
                                                                 reverse=TRUE))

plot_fin_1 <- plot_param[[1]] / plot_param[[2]] / plot_param[[3]]
plot_fin_1 <- grid.arrange(patchworkGrob(plot_fin_1), 
                           left = "Posterior estimate with 95% PIs",
                           bottom = "Transition location")
ggsave("../figures/psplots1constr.pdf", plot_fin_1, height = 8, width = 7)

plot_fin_2 <- plot_param[[4]] / plot_param[[5]] / plot_param[[6]]
plot_fin_2 <- grid.arrange(patchworkGrob(plot_fin_2), 
                           left = "Posterior estimate with 95% PIs",
                           bottom = "Transition location")
ggsave("../figures/psplots2constr.pdf", plot_fin_2, height = 8, width = 7)
```

\begin{figure}[!htb]
\centering
\includegraphics{figures/psplots1constr.pdf}
\caption{Posterior parameter distribution}
\end{figure}
\newpage
\begin{figure}[!htb]
\ContinuedFloat
\captionsetup{list=off,format=cont}
\centering
\includegraphics{figures/psplots2constr.pdf}
\label{fig:fullps1}
\caption{Posterior parameter distribution}
\end{figure}