### 2024-04-27 report/manuscript.Rmd

---
title             : "Ideas cascading into keystrokes -- modelling writing hesitations as finite mixture process"
shorttitle        : "Modelling writing hesitations"

csl               : "apa.csl"

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Mark Torrance"
    affiliation   : "1"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Evgeny Chukharev"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"

  - id            : "2"
    institution   : "Artificial Intelligence Systems Institute, Eindhoven University of Technology, The Netherlands"

  - id            : "3"
    institution   : "Department of English, Iowa State University, Iowa"

abstract: Classical serial models view the process of producing a text as a discrete chain of pauses and writing bursts. In contrast, parallel cascading models of writing assume that planning is not complete at production onset and operates in parallel to writing execution. We implemented these two views in Bayesian statistical models and applied our models to key-stroke logs of 6 data sets from free text production. We reanalysed keystroke intervals at before-sentence transitions, before word transitions and within word transitions. Model comparisons demonstrated strong evidence in favour of the statistical implementation of the cascading view of the serial models across all data sets. Further we found -- across populations -- that transitions across word and even sentence boundaries are often fluent and uninterrupted by planning of upcoming linguistic material; when pausing occurred pauses were longer before sentence than before word which is consistent with the idea larger linguistic edges are associated with higher level planning. Our results cannot be explained by the serial but are in line with the cascading view of writing.

keywords: "Keystroke modelling; finite mixture models; Bayesian models; text composition"

bibliography      : ["references.bib"]

documentclass     : "apa7"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: TRUE

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csquotes          : true

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      comment = NA,
                      cache.extra = R.version,
                      dev = "cairo_pdf",
                      fig.align = 'center', 
                      fig.width = 8, 
                      fig.height = 4.5,
                      width=90)

options(kableExtra.auto_format = FALSE)
dev.args <- list(pdf = list(type = "cairo"))
```

```{r libs, include=FALSE}
library(tidyverse)
library(janitor)
library(brms)
library(papaja)
library(polspline)
library(patchwork)
library(gridExtra)
library(grid)
library(gtable)
library(scales)
library(ggthemes)
source("../scripts/functions.R")
source("scripts/get_pdens_plot.R")

theme_set(theme_few(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

options(scipen = 999)

beta_label <- "Fluent transition\nduration (in msecs)"
delta_label <- "Slowdown for hesitant\ntransitions (in secs)"
theta_label <- "Probability of hesitant\ntransitions"
```

# Introduction

```{r child = "sections/introduction.Rmd"}
```

# Methods

```{r child = "sections/methods.Rmd"}
```

# Results

```{r child = "sections/results.Rmd"}
```

```{=html}
<!-- TODO: 
- address difference between posteriors of shift analysis for constrained and unconstrained model?
- pausing frequency for words and sentences are equal but more importantly pauses occure less than 80% of the time indicating parallel planing.
- would be really good, on some point, to distinguish between open and closed class words
-->
```

# Discussion

```{r child = "sections/discussion.Rmd"}
```

# Conclusion

```{r child = "sections/conclusion.Rmd"}
```

```{=html}
<!--

----
Add to OSF
Work on tutorial and add as link to methods
----

~~~~~
Say how the results don't contradic the reported studies but speaks better to their conclusions 
but the modelling frameword doesn't require us to impose pause thresholds
~~~~~

~~~~
Move from results to discussion?
Open: to what extent are even short latencies longer at sentences. This is obvious for when multiple keystrokes are summed across at sentences boundaries as there are typically three key transitions necessary to start a sentence: (1) .^_ , (2) _^[shift], (3) [shift]^C
We believe the second one is the linguistically meaningful one
I think give Davids papers as examples, if he uses sums instead of means (other studies?)

~~~

neural oscillationsoccur every 153 ms (6.5 hz). Which is kind of the magic number for the short-duration IKI component.

Duprez, J., Stokkermans, M., Drijvers, L., & Cohen, M. X. (2021). Synchronization between keyboard typing and neural oscillations. Journal of Cognitive Neuroscience, 33(5), 887–901. https://doi.org/10.1162/jocn_a_01692

González, N., & Calot, E. P. (2023). Dataset of human-written and synthesized samples of keystroke dynamics features for free-text inputs. Data in Brief, 48, 109125. https://doi.org/10.1016/J.DIB.2023.109125

~~~~
I think the serial account isn't all inconsistent with the mixture model. I don't think the serial account says that pauses must occure on linguistic edges but between bursts which can in principlpe happen anywhere (but why would they be random). Obviously the problem here is that one needs to postulate what a pause is instead of leaving it to the model to work out. The serial account and the cascading account make very similar predictions for inexperienced writers.

Maybe distinguih between a strong version of serial model and a weak version that is basically a chunky cacading model

-->
```

\newpage

# References

\begingroup

\setlength{\parindent}{-0.5in}

\setlength{\leftskip}{0.5in}

<div id = "references"></div>

\endgroup

\newpage

# (APPENDIX) Appendix {.unnumbered}

```{r child = 'appendix/models.Rmd'}
```

\newpage

```{r child = 'appendix/fit2data.Rmd'}
```

\newpage

```{r child = 'appendix/transloc.Rmd'}
```

\newpage

```{r child = 'appendix/shift.Rmd'}
```

\newpage

```{r child = 'appendix/shift_constr.Rmd'}
```

\newpage

```{r child = 'appendix/l2_constr.Rmd'}
```

\newpage

```{r child = 'appendix/masking_constr.Rmd'}
```

\newpage

```{r child = 'appendix/postparam_constr.Rmd'}
```

\newpage

```{r child = 'appendix/prepost_constr.Rmd'}
```

\newpage

```{r child = 'appendix/genre_constr.Rmd'}
```

\newpage

```{r child = 'appendix/transloc_constr.Rmd'}
```

\newpage

```{r child = 'appendix/sim.Rmd'}
```

\newpage

```{r}
#render_appendix("appendix_postparam.Rmd")
#render_appendix("appendix_genre.Rmd")
#render_appendix("appendix_prepost.Rmd")
#render_appendix("appendix_masking.Rmd")
#render_appendix("appendix_l2.Rmd")
```