### 2024-04-27 report/sections/conclusion.Rmd

In contrast to the serial pause-and-burst view, the cascading view emphasis that writers do not necessarily pause to plan the upcoming language unit but plan in parallel to writing execution. Using the Bayesian framework we implemented models of both the serial model of writing and the cascading view. Model comparisons and the inspection of posterior parameter estimates supported with a cascading view of writing but provided strong evidence against the serial view. We found that key transitions across word and even sentences boundaries are frequently fluent and not interrupted but planning of upcoming linguistic information. This pattern was found to be largely consistent for populations with different levels of writing experience and languages (e.g. young / L2 writers, students) and writing tasks (e.g. essays, syntheses).