### 2024-04-27 report/sections/discussion.Rmd

The classical serial view [@flower1980ynamics] characterizes writing as a sequence of planning and execution cycles that results in a writing process that consists of bursts interrupted by pauses. The cascading model of writing, in contrast, captures that planning, at least in competent writers can occur in parallel to production [@olive2014toward]. We used Bayesian statistical models, in particular hierarchical single-distribution and two-distribution mixture models, to directly compare the serial and the cascading model of writing, respectively. On the basis of six datasets with key data from free text production tasks, we presented compelling and consistent statistical evidence in favour of the cascading view.

Patters observed in the parameter estimates are generally similar across data sets (with caveats). Under the cascading view of writing, key transitions that directly precede sentences or words are not necessarily associated with a pause because writers can plan in parallel to writing. On the basis of parameter estimates of the mixture model we tested to what extent larger linguistic edges are associated with the preparation of upcoming planning unit. We found that in the majority of datasets, pauses were as likely to occur before sentences and words, in both experienced and novice writers, while mid-word pauses are generally rare. Also, with few exceptions, pauses were typically longer before sentences than before words (which were longer than mid-word pauses). In other words, when writers do pause, the involvement of higher-level processing at sentence boundaries is typically leading to longer pauses. However, the pausing probability results can only be explained by the cascading view of writing, not by the serial view. This is because our results for the hesitation probability suggest that writers often do not pause before sentences; indeed, hesitation probabilities are largely identical before sentences and before words. This suggest that writers frequently plan the next word and even sentence in parallel to production which mitigates the need to pause to plan up-coming linguistic information.

Interestingly even the youngest writers in our data pool (C2L1) showed evidence of parallel planning expressed in fluent transitions between words and even sentences. Key-transitions were substantially longer compared to all other datasets with a larger slowdown before words. Yet the overall pausing probability showed a patterns comparable to most other data sets. In particular, a larger pause probability was observed for transitions before sentences / words compared to mid-word transitions. This is interesting because, for example, @olive2014toward described a cascading model that is operating in a serial fashion for inexperienced or struggling writers explained on the basis of task demands that reduce the ability to plan in parallel to processing which leads to a serialisation of planning units. However, the results presented in this paper suggest that pausing behaviour in young writers largely mirror results from more experienced writers but pauses were more likely before words and sentences but similarly long indicating a tendency to prepare utterances in planning units possibly smaller than the clause. The main difference for these young writers was that their writing speed (fluent transitions) was substantially slower compared to more experienced writers in the sample. Also for L2 writers (PlanTra) we observe pausing patterns that do not resemble a serialised writing process as indicated by a larger number of fluent transitions across word and sentence boundaries.

We observed two more differences between data sets -- other than the patterns observed for C2L1 summarised above -- that are worth highlighting. These can largely be explained on the basis of differences associated with the population (writing competence / experience) and text genre [see also @conijn2019understanding] that these samples were obtained from: The first exception concern the GUNNEXP2 and SPL2 data sets: pauses were more likely to occur at before-sentence transitions compared to before-word transitions while in the remaining data sets hesitations were equally probably at before-word and before-sentence locations. **TODO: I don't have an explanation for this.** Second, pauses were not always longer before sentence compared to before words compared to mid-word paues; there were two exceptions: (1) no difference was found for before-sentence compared to before-word transitions in C2L1 possibly due to a a word-level planning strategy in this sample of young writers as inexperienced writers tend to use a more localised strategy for text production. (2) pause durations did not vary by transition location in the LIFT dataset although the posterior showed the same general patterns as observed in most other datasets. The absence of a difference could be explained as specific to the synthesis writing task because there is no need for content creation [**REFERENCE?**] but the results for the PlanTra dataset rule out this possibility.

A possible concern with our result -- substantially better predictive performance for two-distributions mixture models -- is that, in principle, as the mixture model has more parameters it might always lead to a better fit. This concern is preempted by using cross-validation techniques for model comparison which is preventing model overfit. In addition we used a simulation approach to compare a single distribution model and a two distributions mixture model. We simulated two data sets that were either based on a single distribution or two-distributions random data generating process. This approach allows us to test the predictive performance of models in a context where we know the true underlying data generating process -- single distribution vs a mixture of two distributions -- and we can test to what extent these models can successfully uncover the true parameter values. Modelling details can be found in Appendix \ref{simulation}. As might be expected, we found, that the two-distribution mixture was successful at uncovering the parameter values of the data generated with a two-distributions mixture models; we observed the same for the single-distribution model of the data generated with a single distribution process. Parameters were not successfully uncovered when we switched model type and dataset. Importantly though, cross-validation did not show a higher predictive performance for the mixture model, compared to the single distribution model, when applied to data based on a single distributions process but a substantially higher predictive performance for data simulated from a two-distributions mixture process. Therefore we can rule out the possibility that the overwhelming statistical support for the two-distributions mixture models can be explained in terms of overfitting.

Throughout this paper we assumed that hesitations are indicators of planning upcoming ideas and encoding of linguistic units. Alternatively at least some if not all pauses in the writing process might be the result of reading or looking back into previously produced text. There is limited research that used eye tracking to investigate regressive gaze behaviour during writing [@alamargot2010using; @beers2010adolescent; @chukharev2019combined; @de2018exploring; @torrancea; @van2010reading]. There are at least four reasons why eye movements away from the inscription point might occure during writing: First, reading is necessary for revision (i.e. writers must check if writing goals have been met). Second, writers might use text to cue or reinstating ideas after interruption of writing process. Third, writers might look for or correct spelling or grammatical errors. We removed transitions that terminated in an editing operation so the pauses we detected are unlikely to reflect revisions of the previous produced text. Fourth, writers might look at the source their are writing from. For two datasets (PlanTra, LIFT) which involved writing from sources, we removed transitions that were made when the sources (rather than the writing document) was in focus.

Frequencies of lookbacks during writing follow a similar pattern as the one we observed for pausing: for example @torrance2016reading reported that lookbacks appear with a frequency of 45% before sentences, 12% before words and 5% within words of which 36% were associated with sustained reading but mostly less patterned forward and backward saccades between words ["hopping", see also @chukharev2019combined]. This is except, lookbacks are more frequent before sentences compared to words while we found that pause frequencies appear equally often before sentences and words. Finally even though reading is likely be explain some proportion of our results it is overall rare, certainly in samples of university students. This can be seen in the result of two datasets in our pool (CATO, GUNNEXP2): these experiments included a manipulation where the produced text was masked (or not). In a reanalysis we found no substantial differences in pausing behaviour for the masked condition; see Appendix \ref{masking-effect-cato-gunnexp2}. This suggest that pausing patterns do not change if reading previous text is not possible.

A practical advantage of the mixture-model approach [also @roeser2021modelling] is that we are not required to stipulate interval thresholds as plausible candidate for lower pause bound to handle the complex distribution of latency from spontaneous text production. Such threshold are widely used by writing researchers, a strategy that was inherited from early research in speech production [for review see @rochester1973significance], with studies dating back to at least the mid 1990s [@foulin1998extent]. There is one central problems with this approach: A threshold requires a definition of what passes as a pause [@wen06; @van2016keystroke], i.e. a pause criterion threshold often set to 2 secs [@chanquoy1996writing; @kaufer1986composing; @sullivan2002self; @wen02] or some other lower bound [@chukharev2014pauses; @connelly2012predicting; @leijten2013keystroke]. Even if these thresholds were adjusted relative to factors such as writing medium, experience of writer, and text location of pause [see e.g. @wen06] they would be arbitrary. For example, a sentence-initial pause of 2 secs has a very different interpretation from a 2 secs pause that occurs before or within a word. Our current understanding of the processes that underlie text production does not provide a strong theoretical basis on which to make this decision. So while a 2 secs threshold undoubtedly captures an interesting distinction -- processing that occurs the range zero to 2 secs is very likely to be qualitatively different from processing that takes more than 2 secs. However the same could be argued for any threshold between perhaps 250 msecs and 10 secs that a researcher might care to choose [@chenu2014interword]. Mixture models provide a principled statistical framework that allows the researcher to model behavioural data that come from a combination of cognitive processes without imposing threshold values [see also @almond2012preliminary; @baaijen2012keystroke; @hall2022constructing; @li2021identifying].