### 2024-05-02 report/manuscript.Rmd
update tutorial final for now

---
title             : "Ideas cascading into keystrokes -- modelling writing hesitations as finite mixture process"
shorttitle        : "Modelling writing hesitations"

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : 
  papaja::apa6_pdf:
    keep_tex: FALSE

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Mark Torrance"
    affiliation   : "1"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Gunn Helen Ofstad"
    affiliation   : "3"
  - name          : "Evgeny Chukharev"
    affiliation   : "4"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"
  - id            : "2"
    institution   : "Artificial Intelligence Systems Institute, Eindhoven University of Technology, The Netherlands"
  - id            : "3"
    institution   : "Educational Sciences and Humanities, University of Stavanger, Norway"
  - id            : "4"
    institution   : "Department of English, Iowa State University, Iowa"

abstract: "Classical serial models view the process of producing a text as a chain of discrete pauses and writing bursts. In contrast, parallel cascading models of writing assume that planning is not complete at production onset and operates in parallel with execution. These two theories make different assumptions about the underlying process that generates inter-keystroke intervals. Across six datasets comprising multi-sentence text composition we instantiated these assumptions as two sets of Bayesian mixed-effects models: (1) single-distribution models as assumed by the serial processing account, and (2) finite two-distribution mixture models as assumed by the parallel processing account. We analysed keystroke intervals at before-sentence, before word, and within word transitions. Model comparisons demonstrated strong evidence in favour of the statistical implementation of the cascading view across all datasets. When pausing occurred, sentence initial pauses were longer than word initial pauses which is consistent with the idea that larger linguistic edges are associated with higher level planning. However, we found -- across populations -- that keystroke latencies at word and even at sentence boundaries were often too rapid to permit planning of what was written next. Our results cannot be explained by the serial but are in line with the cascading, semi-parallel view of writing."

keywords: "Cascading semi-parallel models; serial models; finite mixture models; Bayesian models; text composition; keystrokes"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csquotes          : true

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      comment = NA,
                      cache.extra = R.version,
                      dev = "cairo_pdf",
                      fig.align = 'center', 
                      fig.width = 8, 
                      fig.height = 4.5,
                      width=90)

options(kableExtra.auto_format = FALSE)
dev.args <- list(pdf = list(type = "cairo"))
```

```{r libs, include=FALSE}
library(tidyverse)
library(janitor)
library(brms)
library(papaja)
library(polspline)
library(patchwork)
library(gridExtra)
library(grid)
library(gtable)
library(scales)
library(ggthemes)
source("../scripts/functions.R")
source("scripts/get_pdens_plot.R")

theme_set(theme_few(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

options(scipen = 999)

beta_label <- "Fluent transition\nduration (in msecs)"
delta_label <- "Slowdown for hesitant\ntransitions (in secs)"
theta_label <- "Probability of hesitant\ntransitions"
```

```{r child = "sections/introduction.Rmd"}
```

# Methods

```{r child = "sections/methods.Rmd"}
```

# Results

```{r child = "sections/results.Rmd"}
```

<!-- TODO: 
- would be really good, on some point, to distinguish between open and closed class words
-->

# Discussion

```{r child = "sections/discussion.Rmd"}
```

# Conclusion

```{r child = "sections/conclusion.Rmd"}
```

<!--

Add to OSF
Work on tutorial and add as link to methods
Say how the results don't contradic the reported studies but speaks better to their conclusions

neural oscillationsoccur every 153 ms (6.5 hz). Which is kind of the magic number for the short-duration IKI component.

Duprez, J., Stokkermans, M., Drijvers, L., & Cohen, M. X. (2021). Synchronization between keyboard typing and neural oscillations. Journal of Cognitive Neuroscience, 33(5), 887–901. https://doi.org/10.1162/jocn_a_01692

González, N., & Calot, E. P. (2023). Dataset of human-written and synthesized samples of keystroke dynamics features for free-text inputs. Data in Brief, 48, 109125. https://doi.org/10.1016/J.DIB.2023.109125

-->

\newpage

# References

\begingroup

<!-- \setlength{\parindent}{-0.5in} -->

<!-- \setlength{\leftskip}{0.5in} -->

<div id = "refs"></div>

\endgroup

\newpage

# (APPENDIX) Appendix {.unnumbered}

\newpage

```{r child = 'appendix/models.Rmd'}
```

\newpage

```{r child = 'appendix/fit2data.Rmd'}
```

\newpage

```{r child = 'appendix/transloc.Rmd'}
```

\newpage

```{r child = 'appendix/shift.Rmd'}
```

\newpage

```{r child = 'appendix/shift_constr.Rmd'}
```

\newpage

```{r child = 'appendix/l2_constr.Rmd'}
```

\newpage

```{r child = 'appendix/masking_constr.Rmd'}
```

\newpage

```{r child = 'appendix/postparam_constr.Rmd'}
```

\newpage

```{r child = 'appendix/prepost_constr.Rmd'}
```

\newpage

```{r child = 'appendix/genre_constr.Rmd'}
```

\newpage

```{r child = 'appendix/transloc_constr.Rmd'}
```

\newpage

```{r child = 'appendix/sim.Rmd'}
```

\newpage

```{r}
#render_appendix("appendix_postparam.Rmd")
#render_appendix("appendix_genre.Rmd")
#render_appendix("appendix_prepost.Rmd")
#render_appendix("appendix_masking.Rmd")
#render_appendix("appendix_l2.Rmd")
```