### 2024-05-19 report/appendix/fit2data.Rmd
change tutorial to contrained molg

# Fit to data

Model fit was assessed by testing to what extent simulations obtained from the model fit the observed data. From each model we obtained 50 simulations (predictions); see Figure \ref{fig:prediction}. The model predictions fitted the data substantially better for the mixture models. The worse fit was observed for the single distribution Gaussian model which predicted a larger number of negative keystroke intervals; a log-normal distribution removed these negative predictions.

```{r eval = F}
theme_set(theme_few(base_size = 10) +
              theme(strip.background = element_blank(),
                    legend.position = "top",
                    legend.justification = "right",
                    panel.grid = element_blank(),
                    panel.background = element_rect(fill = "transparent"), # bg of the panel
                    plot.background = element_rect(fill = "transparent", color = NA),
                    axis.text.y = element_blank(),
                    axis.ticks.y = element_blank(),
                    axis.title = element_blank()))

file <- list.files("../../stanout", pattern = "sims", full.names = T, recursive = T)
ds <- str_remove_all(file, "\\.|\\/|stanout|all_sims.csv") %>% 
    str_to_upper() %>% 
    str_replace(., "PLANTRA", "PLanTra") %>%
    str_replace(., "GUNNEXP2", "GE2")

# remove the shift model
ds <- ds[!str_detect(ds, "SHIFT")]

# get all density plots
plots <- map(1:length(ds), ~get_pdens_plot(file[.], max_iki = 2000))

# remove strips after first row
plots[2:length(ds)] <- map(2:length(ds), ~plots[[.]] + theme(strip.text = element_blank()))

# create labels for datasets
labs <- map(ds, ~ggplot(mapping = aes(x = 0, y = 0, label = .x)) + 
  geom_text(angle = 270, size = 2.5) +
  theme_void())

# add names for data sets on the right
plots_labs <- map(1:length(ds), ~plots[[.]] + labs[[.]] + plot_layout(widths = c(.95, .05)))

# combine legends
plots_ledg <- wrap_plots(plots_labs, guides = "collect", ncol = 1)

# add axe titles
plots_fin <- grid.arrange(patchworkGrob(plots_ledg), 
                            left = "Density", bottom = "Interkey interval in secs")

# save plot 
ggsave("../figures/fitplots.pdf", plots_fin, height = 7.25, width = 6.25)

```

```{r prediction, fig.cap="Predicted data (50 simulations in each cell) in red in comparison to the interkey interval data shown in black by model types and datasets. Model fit to the data was order from the best (left) to the weakest (right). For illustration the x-axis was truncated at 2 secs.", eval = T, fig.height=8.25, fig.width=6.75}
knitr::include_graphics("figures/fitplots.pdf")
```