### 2024-05-19 report/appendix/l2_constr.Rmd

# L2 effect (SPL2)

For the SPL2 data we calculated the L2 effect (i.e. the difference between writing in L2 and L1). The results can be found in Table \ref{tab:l2effect}. The results show longer hesitation duration and more pauses across all transition locations when writing in L2. Only within-word transitions did show negligible evidence for an increased hesitation duration in L2.

```{r l2effect, results = 'asis'}
cellmeans <- ps_spl2_both %>% 
  separate(group, into = c("group", "lang"), sep = "; ") %>%
  select(-ds, -group) %>%
  mutate(across(lang, ~str_remove(., "Language: ")),
         across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, data, lang), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = data, 
              values_from = est) %>% 
  mutate(across(c(location, lang), ~ifelse(param == "beta", "overall", .))) %>% 
  unique()

# cell means and data set difference
cellmeans <- cellmeans %>% 
  select(-`SPL2 (_^[shift] + C)`) %>% 
  pivot_wider(names_from = lang, values_from = `SPL2 (_^[shift])`)

file1 <- list.files(path = "../../stanout/spl2", 
                    pattern = "mogbetaco.+.csv$",
                    full.names = T)

tmp <- map_dfr(file1, read_csv) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  mutate(across(lang, recode, "EN" = "L1", "ES" = "L2")) %>% 
  pivot_wider(names_from = lang, values_from = value) %>% 
  unnest(c(L1, L2)) %>% 
  mutate(diff = L2 - L1) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, recode, theta = "prob")) %>% 
  left_join(cellmeans, by = join_by(location, param)) %>% 
  arrange(param, location) %>% 
  select(param, `Transition location` = location, L1, L2,
         "Difference" = mean, BF)

write_csv(tmp, "../tables/l2_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for language effect. Cell means are shown the slowdown for hesitant interkey intervals when writing in L1 and L2 and the probability of hesitant keyintervals. The language difference is shown on log scale (for durations) and logit scale for probability of hesitant key intervals. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          font_size = "small",
          placement = "b",
          stub_indents = list("Hesitation showdown" = 1:3,
                              "Hesitation probability" = 4:6) )

```