### 2024-05-19 report/appendix/prepost.Rmd

# Pre-post test (PLanTra)

The pre-post test effect for the PLanTra dataset is reported in Table \ref{tab:retesteffect}. Transition durations were shorter at before-word locations for the post-test. Evidence was negligible for any other comparisons.

```{r retesteffect, results = 'asis'}
cellmeans <- ps_plantra %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, group, ds), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper, -ds) %>% 
  pivot_wider(names_from = group, values_from = est)

file <- list.files(path = "../../stanout/plantra", 
                   pattern = "mogbetaun.*.csv$",
                   full.names = T)

tmp <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  pivot_wider(names_from = task, values_from = value) %>% 
  unnest(c(`post-test`,`pre-test`)) %>% 
  mutate(diff = `pre-test` - `post-test`) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob"))) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location,
         `Pre-test` = `pre-test`,
         `Post-test` = `post-test`, 
         "Difference" = mean, BF)

write_csv(tmp, "../tables/posttest_effect_cellmeans.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for post-test effect. Cell means are shown for the pre-test and post-test in msecs for fluent interkey intervals, the slowdown for long interkey intervals and the probability of hesitant transitions. The effect for post-test is shown on log scale (for interkey intervals) and logit scale for probability of hesitant interkey intervals. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Fluent interkey intervals" = 1:3, 
            "Hesitation duration" = 4:6,
            "Hesitation probability" = 7:9))

```