### 2024-05-19 report/appendix/transloc_constr.Rmd

# Transition location effect

```{r loceffectcon, results='asis'}

files <- list.files(str_c("../../stanout/", 
                          c("lift", "spl2", "gunnexp2", "plantra", "cato", "c2l1")), pattern = "mogbetacon.+.csv", full.names = T) # "spl2_shift",

ps_loc_diffs <- purrr::map_dfr(files, ~read_csv(.x) %>% 
                                 mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode(., 
                              "cato" = "*CATO* (non-dyslexic\nunmasked)",
                              "spl2" = "*SPL2* (L1)", 
                              "gunnexp2" = "*GE2* (unmasked)",
                              #                "spl2_shift" = "*SPL2* (L1; shift + C)", 
                              "plantra" = "*PLanTra*",
                              "lift" = "*LIFT*",
                              "c2l1" = "*C2L1*"))) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked"),
         !(str_detect(data, "GE2") & xn == "masked"),
         param %in% c("delta", "theta")) %>%
  select(param, location, data, value) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, .))) %>% 
  pivot_wider(names_from = location, values_from = value) %>%
  unnest(-c(data, param)) %>% 
  mutate(diff.1 = `before sentence` - `before word`,
         diff.2 = `before word` - `within word`) %>% 
  summarise(across(starts_with("diff"), 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper, 
                        BF = BF)),
            .by = c(data, param))

tmp <- ps_loc_diffs %>% 
  pivot_longer(starts_with("diff"), names_to = c("diffid", ".value"), names_sep = "_") %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2))))) %>% 
  select(-lower, -upper) %>% 
  pivot_wider(names_from = param, values_from = c(mean, BF)) %>% 
  mutate(across(diffid, ~recode(., diff.1 = "before sentence vs word",
                                diff.2 = "before vs within word"))) %>% 
  select(data, diffid, ends_with("delta"), ends_with("theta")) %>% 
  arrange(data)

write_csv(tmp, "../tables/location_effect_constr_alldata.csv")

apa_table(tmp[,-1],
          align =c("l", rep("r", 4)), 
          longtable = T,
          escape = FALSE,
          row.names = T,
          font_size = "footnotesize",
          col.names = c("Comparison", rep(c("Est. [95\\% PIs]", "BF"), 2)),
          stub_indents = list(`\\textbf{C2L1}` = 1:2, 
                              `\\textbf{CATO (non-dyslexic, unmasked)}`= 3:4,
                              `\\textbf{GE2 (unmasked)}` = 5:6,
                              `\\textbf{LIFT}` = 7:8,
                              `\\textbf{PLanTra}` = 9:10,
                              `\\textbf{SPL2 (L1)}` = 11:12),
          #                              `\\textbf{SPL2 (L1; shift + C)}` = 13:14),
          col_spanners = list(" " = c(1),
                          #    "Fluent transitions" = c(2, 3), 
                              "Hesitation slowdown" = c(2, 3), 
                              "Hesitation probability" = c(4, 5)),
          note = "PI = probability intervals. BF = evidence in favour of the alternative hypothesis over the null hypothesis.",
          caption = "Effect of transition location on interkey intervals. Differences are shown on log scale (for durations) and logit scale for hesitation probability. 95\\% PIs in brackets.")

```