### 2024-05-19 report/manuscript.Rmd

---
title             : "Ideas cascade in parallel into keystrokes -- modelling writing hesitations as finite mixture process"
shorttitle        : "Ideas cascade in parallel into keystrokes"

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : 
  papaja::apa6_pdf:
    keep_tex: FALSE

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Mark Torrance"
    affiliation   : "1"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Gunn Helen Ofstad"
    affiliation   : "3"
  - name          : "Evgeny Chukharev"
    affiliation   : "4"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"
  - id            : "2"
    institution   : "Artificial Intelligence Systems Institute, Eindhoven University of Technology, The Netherlands"
  - id            : "3"
    institution   : "Educational Sciences and Humanities, University of Stavanger, Norway"
  - id            : "4"
    institution   : "Department of English, Iowa State University, Iowa"

abstract: "Classical serial models view the process of producing a text as a chain of discrete pauses, during which the next span of text is planned, and bursts of activity during which this text is output onto the page or computer screen. In contrast, semi-parallel models of writing assume that planning is not complete at production onset and operates in parallel with execution. Across six datasets from participants composing multi-sentence texts we instantiated these assumptions in two sets of Bayesian mixed effects models: (1) single-distribution models as assumed by the serial processing account, and (2) a finite (two-distribution) mixture model as assumed by the parallel processing account. We analysed interkey intervals at before-sentence, before word, and within word transitions. Model comparisons demonstrated strong evidence in favour of the parallel view across all datasets. When pausing occurred, sentence initial pauses were longer than word initial pauses which is consistent with the idea that larger linguistic edges are associated with higher level planning. However, we found -- across populations -- that interkey intervals at word and even at sentence boundaries were often too rapid to permit planning of what was written next. Our results cannot be explained by the serial processing but are in line with the semi-parallel view of writing."

keywords: "Cascading semi-parallel models; serial models; finite mixture models; Bayesian models; text composition; keystrokes"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csquotes          : true

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      comment = NA,
                      cache.extra = R.version,
                      dev = "cairo_pdf",
                      fig.align = 'center', 
                      fig.width = 8, 
                      fig.height = 4.5,
                      width=90)

options(kableExtra.auto_format = FALSE)
dev.args <- list(pdf = list(type = "cairo"))
```

```{r libs, include=FALSE}
library(tidyverse)
library(janitor)
library(brms)
library(papaja)
library(polspline)
library(patchwork)
library(gridExtra)
library(grid)
library(gtable)
library(scales)
library(ggthemes)
source("../scripts/functions.R")
source("scripts/get_pdens_plot.R")

theme_set(theme_few(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

options(scipen = 999)

beta_label <- "Fluent interkey\nintervals (in msecs)"
delta_label <- "Slowdown for hesitant\ninterkey intervals (in secs)"
theta_label <- "Probability of hesitant\ninterkey intervals"
```

```{r child = "sections/introduction.Rmd"}
```

# Methods

```{r child = "sections/methods.Rmd"}
```

# Results

```{r child = "sections/results.Rmd"}
```

<!-- TODO: 
- would be really good, on some point, to distinguish between open and closed class words
-->

# Discussion

```{r child = "sections/discussion.Rmd"}
```

# Conclusion

```{r child = "sections/conclusion.Rmd"}
```

<!--

Add to OSF
Work on tutorial and add as link to methods
Say how the results don't contradic the reported studies but speaks better to their conclusions

neural oscillationsoccur every 153 ms (6.5 hz). Which is kind of the magic number for the short-duration IKI component.

Duprez, J., Stokkermans, M., Drijvers, L., & Cohen, M. X. (2021). Synchronization between keyboard typing and neural oscillations. Journal of Cognitive Neuroscience, 33(5), 887–901. https://doi.org/10.1162/jocn_a_01692

González, N., & Calot, E. P. (2023). Dataset of human-written and synthesized samples of keystroke dynamics features for free-text inputs. Data in Brief, 48, 109125. https://doi.org/10.1016/J.DIB.2023.109125

-->

\newpage

# References

\begingroup

<!-- \setlength{\parindent}{-0.5in} -->

<!-- \setlength{\leftskip}{0.5in} -->

<div id = "refs"></div>

\endgroup

\newpage

# (APPENDIX) Appendix {.unnumbered}

\newpage

```{r child = 'appendix/models.Rmd'}
```

\newpage

```{r child = 'appendix/fit2data.Rmd'}
```

\newpage

```{r child = 'appendix/transloc.Rmd', eval=F}
```

\newpage

```{r child = 'appendix/shift.Rmd', eval = F}
```

\newpage

```{r child = 'appendix/shift_constr.Rmd'}
```

\newpage

```{r child = 'appendix/l2_constr.Rmd'}
```

\newpage

```{r child = 'appendix/masking_constr.Rmd'}
```

\newpage

```{r child = 'appendix/postparam_constr.Rmd'}
```

\newpage

```{r child = 'appendix/prepost_constr.Rmd'}
```

\newpage

```{r child = 'appendix/genre_constr.Rmd'}
```

\newpage

```{r child = 'appendix/transloc_constr.Rmd'}
```

\newpage

```{r child = 'appendix/sim.Rmd'}
```

\newpage

```{r}
#render_appendix("appendix_postparam.Rmd")
#render_appendix("appendix_genre.Rmd")
#render_appendix("appendix_prepost.Rmd")
#render_appendix("appendix_masking.Rmd")
#render_appendix("appendix_l2.Rmd")
```