### 2024-05-28 report/manuscript.Rmd
Submitted paper

---
title             : "Typing in tandem: finite mixture models reveal fundamental parallelism in multi-sentence text composition"
shorttitle        : "Typing in tandem"

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : 
  papaja::apa6_pdf:
    keep_tex: TRUE

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Evgeny Chukharev"
    affiliation   : "3"
  - name          : "Gunn Helen Ofstad"
    affiliation   : "4"
  - name          : "Mark Torrance"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"
  - id            : "2"
    institution   : "Human-Technology Interaction Group, Eindhoven University of Technology, The Netherlands"
  - id            : "3"
    institution   : "Department of English, Iowa State University, Iowa"
  - id            : "4"
    institution   : "Educational Sciences and Humanities, University of Stavanger, Norway"

# We are grateful for all authors who made their available (to us), in particular Nina Vandermeulen, Alessandra Rossetti, and colleagues. This work was supported by the National Science Foundation under Grant No. 2016868: "ProWrite: Biometric feedback for improving college students' writing processes." and UKRI ESRC under Grant No. ES/W011832/1: "Can you use it in a sentence?: Establishing how word-production difficulties shape text formation."
# 
# Data and analysis scripts written in R and Stan can be found on OSF (https://osf.io/z65dw/; Roeser et al., 2024). For a tutorial on Bayesian mixture-model analysis in the context of keystroke data see https://rpubs.com/jensroes/mixture-models-tutorial.

abstract: 'Classical serial models view the process of producing a text as a chain of discrete pauses, during which the next span of text is planned, and bursts of activity during which this text is output onto the page or computer screen. In contrast, semi-parallel models of writing assume that planning is not complete at production onset and operates in parallel with execution. Across six datasets from participants composing multi-sentence texts we instantiated these assumptions in two sets of Bayesian mixed effects models: (1) single-distribution models as assumed by the serial processing account, and (2) a finite (two-distribution) mixture model as assumed by the parallel processing account. We analysed interkey intervals at before-sentence, before word, and within word transitions. Model comparisons demonstrated strong evidence in favour of the parallel view across all datasets. When pausing occurred, sentence initial pauses were longer than word initial pauses which is consistent with the idea that larger linguistic edges are associated with higher level planning. However, we found -- across populations -- that interkey intervals at word and even at sentence boundaries were often too rapid to permit planning of what was written next. Our results cannot be explained by the serial processing but are in line with the semi-parallel view of writing.'

keywords: "Cascade; parallel models; serial models; finite mixture models; Bayesian models; text composition; keystrokes"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csquotes          : true

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      comment = NA,
                      cache.extra = R.version,
                      dev = "cairo_pdf",
                      fig.align = 'center', 
                      fig.width = 8, 
                      fig.height = 4.5,
                      width=90)

options(kableExtra.auto_format = FALSE)
dev.args <- list(pdf = list(type = "cairo"))
```

```{r libs, include=FALSE}
library(tidyverse)
library(janitor)
library(brms)
library(papaja)
library(polspline)
library(patchwork)
library(gridExtra)
library(grid)
library(gtable)
library(scales)
library(ggthemes)
source("../scripts/functions.R")
source("scripts/get_pdens_plot.R")

theme_set(theme_few(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))

options(scipen = 999)

beta_label <- "Fluent interkey\nintervals (in msecs)"
delta_label <- "Slowdown for hesitant\ninterkey intervals (in secs)"
theta_label <- "Probability of hesitant\ninterkey intervals"
```

```{r child = "sections/introduction.Rmd"}
```

# Methods

```{r child = "sections/methods.Rmd"}
```

# Results

```{r child = "sections/results.Rmd"}
```

<!-- TODO: 
- would be really good, on some point, to distinguish between open and closed class words
-->

# Discussion

```{r child = "sections/discussion.Rmd"}
```

# Conclusion

```{r child = "sections/conclusion.Rmd"}
```

<!--

Add to OSF
Work on tutorial and add as link to methods
Say how the results don't contradic the reported studies but speaks better to their conclusions

neural oscillationsoccur every 153 ms (6.5 hz). Which is kind of the magic number for the short-duration IKI component.

Duprez, J., Stokkermans, M., Drijvers, L., & Cohen, M. X. (2021). Synchronization between keyboard typing and neural oscillations. Journal of Cognitive Neuroscience, 33(5), 887–901. https://doi.org/10.1162/jocn_a_01692

González, N., & Calot, E. P. (2023). Dataset of human-written and synthesized samples of keystroke dynamics features for free-text inputs. Data in Brief, 48, 109125. https://doi.org/10.1016/J.DIB.2023.109125

-->

\newpage

# References

\begingroup

<!-- \setlength{\parindent}{-0.5in} -->

<!-- \setlength{\leftskip}{0.5in} -->

<div id = "refs"></div>

\endgroup

\newpage

# (APPENDIX) Appendix {.unnumbered}

\newpage

```{r child = 'appendix/models.Rmd'}
```

\newpage

```{r child = 'appendix/fit2data.Rmd'}
```

\newpage

```{r child = 'appendix/transloc.Rmd', eval=F}
```

\newpage

```{r child = 'appendix/shift.Rmd', eval = F}
```

\newpage

```{r child = 'appendix/shift_constr.Rmd'}
```

\newpage

```{r child = 'appendix/l2_constr.Rmd'}
```

\newpage

```{r child = 'appendix/masking_constr.Rmd'}
```

\newpage

```{r child = 'appendix/postparam_constr.Rmd'}
```

\newpage

```{r child = 'appendix/prepost_constr.Rmd'}
```

\newpage

```{r child = 'appendix/genre_constr.Rmd'}
```

\newpage

```{r child = 'appendix/transloc_constr.Rmd'}
```

\newpage

```{r child = 'appendix/sim.Rmd'}
```

\newpage

```{r}
#render_appendix("appendix_postparam.Rmd")
#render_appendix("appendix_genre.Rmd")
#render_appendix("appendix_prepost.Rmd")
#render_appendix("appendix_masking.Rmd")
#render_appendix("appendix_l2.Rmd")
```