### 2024-05-28 report/sections/conclusion.Rmd

The parallel view of writing, in contrast to the serial view, claims that writers do not necessarily pause to plan the upcoming language unit but plan in parallel with output. In this paper we have demonstrated, across six datasets, that statistical models premised on a parallel-processing understanding of planning in written composition outperform models premised on serial processing. We also found that interkey intervals across word and even sentence boundaries are frequently at timescales that are not consistent with planning of upcoming linguistic information. This pattern was found to be largely consistent for populations with different levels of writing experience and languages (e.g. young / L2 writers, students) and across different composition tasks. Our argument that written composition -- a very common form of language production -- is a largely parallel process is probably uncontroversial. However, the findings we presented in this paper constitute the first robust evidence in support of this claim.