### 2024-05-28 report/sections/discussion.Rmd

The classical serial view [@flower1980ynamics] characterises composing spontaneous multi-sentence text as a sequence of planning and execution cycles. Text is therefore produced as a series of production bursts bounded by pauses. The parallel model of writing, in contrast, sees  planning, at least in competent writers, as occurring to a large extent in parallel with output [@olive2014toward]. The implication of this for observation of writing timecourse -- of inter-key intervals as writers compose text -- is that pause duration doesn’t necessarily correlate with planning necessary for next output. We implemented  these two views in Bayesian statistical models, hierarchical single-distribution models capturing the serial view, and a two-distribution mixture model capturing predictions about patterning of interkey intervals under the assumption of semi-parallel upstream processing. Fitting these models with six sets of keystroke data from free text production tasks -- with a total of 953 participants producing 5,223,280 interkey intervals -- we have found compelling and consistent statistical evidence in favour of the parallel view. In all cases, the predictive performance of the two-distribution mixture-models (consistent with the parallel view of writing) outperformed the single-distribution models.

The two distribution model assumes two data generating processes, one associated with the immediate demands of motor planning required for the production of a keystroke, and one that captures all upstream processing from message to abstract letter representation. A direct implication of parallel, cascaded processing is that the distribution over interkey intervals is not additive -- it is not a combination of time to motor plan plus time for upstream processing. The interval between keypresses is *either* determined by motor execution *or* by processing that occurred upstream of motor execution (in cases where provision of output from the upstream processes lags behind the rate at which it can be output by the fingers). Under a serial account -- where all planning necessary for production of next-text occurs in the interkey interval immediately before output of -- interkey interval will be the sum of times for each processing step required for generation of that text. Both accounts predict longer mean interkey intervals at higher-order text edges (interkey intervals greater before sentences than before mid-sentence words, and before mid-sentence words than before mid-word characters). This has been consistently observed in a number of previous studies [@torrance2016adolescent; @wen02; @mohsen2021second; @miller2006pausing; @xu2017analyzing; @miller2008psycholinguistic;@deane2015exploring]. However, the serial account does not permit the possibility of a distribution of short interkey intervals at higher-order edges: It strongly predicts that *all* sentence boundaries will be associated with interkey intervals that are longer than those required just for the motor planning of the next character. Under the parallel-processing account it is very likely that some -- perhaps a large majority -- of interkey intervals before sentences will be longer. However there will be a distinct subset that is drawn from a distribution associated just with interkey intervals necessary for motor execution.

This is what we found. At all linguistic edges we found a distribution of interkey intervals at durations to motor execution at durations substantially below a threshold that might plausibly be necessary for, for example, lexical / orthographic retrieval [@grudin1982digraph; @ostry1983determinants; @terzuolo1980determinants; @waes2019]. For mid-word interkey intervals, as might be expected, the mixture model predicted that the vast majority of interkey intervals fall within this distribution. Across the six datasets that we sampled, the central tendency for the duration of mid-word interkey intervals was broadly similar to mean mid-word interkey intervals observed in previous studies. Importantly, however, these short interkey intervals were also present at higher-order text edges. Across all datasets we found, at minimum, a non-trivial minority of sentence-initial interkey intervals were sampled from the short distribution (i.e. with durations that could only be associated with motor execution). In two cases -- for writers sampled in the PLanTra and Lift datasets -- the majority of interkey intervals fell within this distribution. Thus, although as might be expected there was a considerably increased tendency for longer interkey intervals at higher order boundaries: Planning associated with planning a new sentence is, probabilistically, much more likely to spill beyond output of the previous sentence and therefore affect duration of the sentence-initial interkey interval. However the presence of a non-trivial proportion of interkey intervals with a duration that is inconsistent with anything other than motor execution can only be explained with recourse to theories that permit preparation in parallel with preceding output.

Our results showed that the hesitation probabilities are largely identical before sentences and before words, which suggests that writers frequently plan the next word and even sentence in parallel to production -- consistent with the parallel view of writing -- which mitigates the need to pause to plan up-coming linguistic information.

This parallelism was evident even in the youngest writers in our data pool (C2L1; mean age: 11.8 years). As might be expected for children who compose predominantly by handwriting and whose typing skills will still be developing, motor execution rate (mean duration of fluent transitions) was substantially slower compared to the more experienced typists in other datasets. In all other respects, however, findings for this dataset were similar to that of other datasets and consistent with parallel processing. A two-distribution mixture model provided best fit to these data. At both before-word and before-sentence locations a non-trivial number of interkey intervals were within the short distribution - i.e. at durations too short to permit lexical retrieval or syntactic planning.    Probability of longer intervals was greater before-sentence and before-word transitions compared to mid-word transitions. This is interesting because, for example, @olive2014toward described a parallel model that operates in a serial fashion for inexperienced or struggling writers, explained on the basis of task demands that reduce the ability to plan in parallel to processing which leads to a serialisation of planning units. However, the results presented in this paper suggest that pausing behaviour in young writers largely mirror results from more experienced writers: Pauses were both equally long and likely at before-word and before-sentence transitions but more likely compared to mid-word transitions. Also for L2 writers (PLanTra) we observe pausing patterns that do not resemble a serialised writing process as evidenced by a large number of fluent transitions at word and sentence boundaries.

We observed two more differences among datasets that are worth highlighting. Writers sampled in the GE2 and SPL2 datasets were more likely to pause before sentences than before words [see also @medimorec2017pauses; @wen02]. In the remaining datasets from writers who were either younger (and not educationally-selected) or writing in second language,  hesitation was equally probable pre-word and pre-sentence. This is most readily attributed simply to increased demands of lexical and / or orthographic processing in these samples.

 Second, pauses at before-sentence locations were not always longer than before words or mid-word; specifically there were two exceptions: (1) sentence-initial hesitations were not longer than word-initial hesitation for the C2L1 dataset possibly because these sample of young inexperienced writers tend to use a more localised word-level planning strategy for text production. (2) pause durations did not vary across transition locations in the LIFT dataset. The absence of a difference could be explained as specific to the writing task because, in synthesis, there is less need for planning novel contents and text structure: the structure of one the source texts might have been used during synthesis [“structural isomorphism”, @hidi1986producing]. However, the results of the PLanTra dataset rule out this possibility: there was evidence for longer hesitations at larger linguistic edges for a text simplification task that also does not require planning of contents.

 We are, therefore, interpreting the superior predictive performance of the two-distribution mixture model as strong evidence for parallelism in written composition. An alternative, more prosaic explanation might be that two-distribution models outperform single distributions simply because they estimate more parameters; two distributions will always provide better fit than one distribution models due just to overfitting. There are two related reasons not to prefer this explanation. First, the cross-validation methods that we used to compare model predictive performance favour more parsimonious models and are therefore robust to overfitting. Second, in the analysis described in Appendix \ref{simulation} we directly tested the hypothesis that modelling data randomly sampled from a single log-normal distribution as a two-distribution mixture will provide better fit than modelling as a single distribution. We found that the two-distribution mixture model was successful at uncovering the parameter values of the data generated with a two-distributions mixture process; we observed the same for the single-distribution model applied to the data generated with a single distribution process. Parameters were not successfully uncovered when we switched model type and dataset. Cross-validation rendered a substantially higher predictive performance for the mixture model compared to the single-distribution model for data sampled from a two-distributions mixture process, but not when applied to data based on a single-distribution process. Therefore we can rule out the possibility that the statistical support for the two-distributions mixture-models can be explained on the basis of their larger number of parameters (i.e.model overfitting).

Throughout this paper we have been attributing longer (second-distribution) interkey intervals to spill-over from upstream pre-motor processing that could not be completed in its entirety in parallel with previous output. These longer inter-key intervals may also, on occasion, be associated with activity that is initiated after the preceding process and is entirely bounded by the duration of the pause. We are thinking here specifically of the tendency, particularly at higher order text boundaries, for writers to glance back into their existing text [@alamargot2010using; @beers2010adolescent; @chukharev2019combined; @de2018exploring; @torrancea; @van2010reading]. @torrance2016reading found that lookbacks appear with a frequency of 45% before sentences, 12% before words and 5% within words of which 36% were associated with sustained reading but mostly less patterned forward and backward saccades between words ["hopping", see also @chukharev2019combined]. There are broadly two reasons why writers might look back into their existing text.^[Eye movement between key presses in writing-from-sources tasks will also be associated with reading source texts. Note, however, that for  two datasets (PLanTra, LIFT) that involved writing from sources, we removed transitions that were made when the source (rather than the writing document) was in focus.]  Lookback may be associated with monitoring just-produced text for errors. In the present study we excluded interkey intervals that preceded deletion or cursor movement, and also intervals at timescales that might be associated with strategic reviewing of large spans of text (e.g., reading text-already-written from top to bottom). However it may be that some longer intervals were associated with monitoring that failed to find errors that needed correcting. A second function for lookback may be to support planning next-text [@johansson2010looking; @hayes2013new; @torrance2016reading]. The preceding text, and particularly the preceding sentence, is a rich source of cues to support retrieval of both content and lexical items necessary to prepare what to write next. This will become particularly important where upstream processing lags behind the rate at which it can be output, and one or more of the buffers necessary to resolve this “temporal friction” [@van1991handwriting] becomes overloaded. Comparison of conditions unmasked and masked text conditions in the CATO and GE2 datasets -- conditions in which writers could and could not read the text that they had already produced -- suggested that, if anything, proportion of longer interkey intervals at sentence boundaries was greater when text was masked (details in Appendix \ref{masking-effect-cato-ge2}). This finding is consistent both with lookback being only a minor contributor to interkey interval and, where it does occur, serving a rapid memory-refresh function. There is also some evidence, at least in the context of handwritten production, that eye movements back into already-written text themselves occasionally occur in parallel with output [@Alamargot2007].

The findings we report in this paper are, we argue, the first robust evidence for parallelism in written composition of multi-sentence texts. The conclusion that text production processes are  fundamentally parallel will not come as a surprise to most researchers. Parallel processing has a long history in theories of speech production [@gar75; @levelt1989speaking; @dell1986; @chang2006; @dell1999connectionist; @dell1992stages]. The transitory nature of activation and storage during fluent spoken production prevent, for example, generating phonological code a long time before it is required to drive articulation [@christiansen2016now]. Spoken production is “just-in-time”. Perhaps because, at least in principle, there is not the same push to fluency in written production -- in principle writers can pause at any point with no detriment to communication -- writing extended text has tended to be seen as fundamentally serial, with pauses to prepare next-text followed by bursts of output during which this is executed [@flower1980ynamics]. Our findings suggest that this is not the case. The language production processes that underlie text production follow the same just-in-time principle as speech.

<!-- A practical advantage of the mixture-model approach [also @roeser2021modelling] is that we are not required to stipulate interval thresholds as plausible candidates for lower pause bound to handle the complex distribution of latency from spontaneous text production. Such thresholds are widely used by writing researchers, a strategy that was inherited from early research in speech production [for review see @rochester1973significance], with studies dating back to at least the mid 1990s [@foulin1998extent]. There is one central problems with this approach: A threshold requires a definition of what passes as a pause [@wen06;@van2016keystroke], i.e. a pause criterion threshold often set to 2 secs [@chanquoy1996writing; @kaufer1986composing; @sullivan2002self; @wen02] or some other lower bound [@chukharev2014pauses; @connelly2012predicting; @leijten2013keystroke]. Even if these thresholds were adjusted relative to factors such as writing medium, experience of writer, and text location of pause [see e.g. @wen06] they would be arbitrary. For example, a sentence-initial pause of 2 secs has a very different interpretation from a 2 secs pause that occurs before or within a word. Our current understanding of the processes that underlie text production does not provide a strong theoretical basis on which to make this decision. So while a 2 secs threshold undoubtedly captures an interesting distinction -- processing that occurs in the range zero to 2 secs is very likely to be qualitatively different from processing that takes more than 2 secs. However the same could be argued for any threshold between perhaps 250 msecs and 10 secs that a researcher might care to choose [@chenu2014interword]. Mixture models provide a principled statistical framework that allows the researcher to model behavioural data that come from a combination of cognitive processes without imposing threshold values [see also @almond2012preliminary; @baaijen2012keystroke; @hall2022constructing; @li2021identifying]. -->