### 2024-06-13 slides/sections/datasets.Rmd

## Data sets

```{r}
tibble(Dataset = c("C2L1", "CATO", "SPL2", "PLanTra", "LIFT", "GE2"),
                Source = c("@ronneberg2022process", # "Rønneberg et al. (2022)",
                            "@torrance2016adolescent", # "Torrance et al. (2016)",
                           "@torrancea",#  "Torrance et al. (n.d.)",
                          "@rossetti2022text",#   "Rossetti and Van Waes (2022b)",
                          "@vandermeulen2020mapping",#  "Vandermeulen et al. (2020)", 
                           "@torranceb"), # "Torrance and Ofstad (n.d)"),
                Keylogger = c("EyeWrite", 
                           "EyeWrite", 
                           "CyWrite", 
                           "InputLog", 
                           "InputLog",
                           "EyeWrite"),
                Task = c("Argumentative", 
                                 "Expository",
                                 "Argumentative",
                                 "Text simplification",
                                 "Synthesis", 
                                 "Argumentative"),
                 N = as.character(c(126, 26*2, 39, 47, 658, 45)),
#                n_texts = c(1, 2, 2, 2, NA),
#                n_sentences = c(),
#                n_words = c(),
                 Age = c(11.8, 16.9, 20.6, 23, 17.0, 19),
                 Sample = c("6th graders", "Secondary school students (dyslexic, non dyslexic)", "Undergraduate students", "Master students", "Secondary school students", "Undergraduate students"),
                 Country = c("Norway", "Norway", "USA", "Belgium", "The Netherlands", "UK"),
                 Language = c( "Norwegian", 
                               "Norwegian", 
                               "English", 
                               "English (L2)", 
                               "Dutch",
                               "English")) %>% 
  arrange(Dataset) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 12) %>% 
  row_spec(0:6, extra_css = "line-height: 2.25;")

```