### 2024-06-13 slides/sections/discussion.Rmd

## Planning text unfolds in parallel to production!

- Writers do not necessarily pause but plan utterances in parallel to writing.
- Evidenced by 
  - stronger predictive performance of mixture models [@roeser2021modelling].
  - pauses are generally equally likely before before sentences compare to before words (but often longer).
- Largely consistent for populations with different levels of writing experience, languages (e.g. young / L2 writers, students), composition tasks. 
- First robust evidence that written composition is largely a parallel process.