### 2024-06-13 slides/sections/implementation.Rmd

## Implentation

- Bayesian inference [e.g. @lambert2018student]
- Random effects for participants [see @roeser2021modelling] with code based on @vasishth2017
- Stan [@carpenter2016stan] using `rstan` [@rstan] to interface with $R$
- Leave-one-out cross-validation estimated as the sum of the expected log predictive density $\widehat{elpd}$ and compared by the difference between models $\Delta\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical]
- Difference between models was summarised as $\mid\frac{\Delta\widehat{elpd}}{\text{SE}}\mid$, the *z*-score of the difference between models [@sivula2020uncertainty].