### 2024-06-13 slides/sections/locationeffects.Rmd

## Mixture distributions

```{r mixmodel, fig.cap='', fig.height=12, fig.width=10, message=FALSE}
# Data grid
data <- tibble(x = c(0, 100))

# plotting details
colours <- c("Fluent" = "#000000", "Hesitant" = "#E69F00")
ylims <- c(0, .7)
xbreaks <- log(c(1, 10, 100, 1000, 10000))
xlims <- log(c(10, 50000))
subtitle_bs <- "Before sentence"
subtitle_bw <- "Before word"
subtitle_ww <- "Within word"

# get files
files <- list.files("../../stanout", pattern = "mog_constr_posterior_for_densityplot.csv", recursive = T, full.names = T)

# load all posteriors
ps_all <- map_dfr(files, ~read_csv(.) %>% mutate(dataset = .x)) %>% 
  mutate(dataname = case_when(str_detect(dataset, "spl2") ~ "SPL2",
                              str_detect(dataset, "lift") ~ "LIFT",
                              str_detect(dataset, "plantra") ~ "PLanTra",
                              str_detect(dataset, "gunnexp2") ~ "GE2",
                              str_detect(dataset, "cato") ~ "CATO",
                              str_detect(dataset, "c2l1") ~ "C2L1"))

# plot counter
plot_id <- 0

for(i in files){

  if(plot_id == 0){
    # template for output
    plot_bs <- list()
    plot_bw <- list()
    plot_ww <- list()
  }

  plot_id <- plot_id + 1

  # select dataset
  ps <- filter(ps_all, dataset == i)

  # add data name
  data <- mutate(data, dataname = unique(pull(ps, dataname)))

  # Extract parameters
  alpha <- filter(ps, param == 'beta') %>% pull(est)
  alpha2_bs <- filter(ps, param == 'delta', loc == 'before sentence') %>% pull(est)
  alpha2_bw <- filter(ps, param == 'delta', loc == 'before word') %>% pull(est)
  alpha2_ww <- filter(ps, param == 'delta', loc == 'within word') %>% pull(est)

  theta_bs <- filter(ps, param == 'prob',  loc == 'before sentence') %>% pull(est)
  theta_bw <- filter(ps, param == 'prob',  loc == 'before word') %>% pull(est)
  theta_ww <- filter(ps, param == 'prob',  loc == 'within word') %>% pull(est)

  sigma <- filter(ps, param == 'sigma') %>% pull(est)
  sigmap_bs <- filter(ps, str_detect(param, 'sigma'), loc %in% c('overall', 'before sentence')) %>% pull(est) %>% sum()
  sigmap_bw <- filter(ps, str_detect(param, 'sigma'), loc %in% c('overall', 'before word')) %>% pull(est) %>% sum()
  sigmap_ww <- filter(ps, str_detect(param, 'sigma'), loc %in% c('overall', 'within word')) %>% pull(est) %>% sum()

  # Create density plot
  plot_bs[[plot_id]] <- ggplot(data, aes(x)) +
    stat_function(aes(colour = "Fluent"),
                  geom = "line", 
                  linewidth = .25,
                  fun = plot_mix_comps,
                  args = list(alpha, sigma, lam = 1 - theta_bs)) +
    stat_function(aes(colour = "Hesitant"),
                  geom = "line", 
                  linewidth = .25,
                  fun = plot_mix_comps, 
                  args = list(alpha2_bs, sigmap_bs, lam = theta_bs)) +
    stat_function(aes(fill = "Fluent"),
                  geom = "area", 
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha, sigma, lam = 1 - theta_bs)) +
    stat_function(aes(fill = "Hesitant"),
                  geom = "area", 
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha2_bs, sigmap_bs, lam = theta_bs)) +
    scale_x_continuous(labels = ~exp(.) / 1000, 
                       breaks = xbreaks, 
                       limits = xlims) +
    scale_y_continuous(limits = ylims) +
    scale_colour_manual(values = colours) +
    scale_fill_manual(values = colours) +
    labs(colour = "Mixture component",
         fill = "Mixture component")

  plot_bw[[plot_id]] <- ggplot(data, aes(x)) +
    stat_function(aes(colour = "Fluent"),
                  linewidth = .25,
                  geom = "line", 
                  fun = plot_mix_comps,
                  args = list(alpha, sigma, lam = 1 - theta_bw)) +
    stat_function(aes(colour = "Hesitant"),
                  geom = "line", 
                  linewidth = .25,
                  fun = plot_mix_comps, 
                  args = list(alpha2_bw, sigmap_bw, lam = theta_bw)) +
    stat_function(aes(fill = "Fluent"),
                  geom = "area", 
                  linewidth = .25,
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha, sigma, lam = 1 - theta_bw)) +
    stat_function(aes(fill = "Hesitant"),
                  geom = "area", 
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha2_bw, sigmap_bw, lam = theta_bw)) +
    scale_x_continuous(labels = ~exp(.) / 1000, 
                       breaks = xbreaks, 
                       limits = xlims) +
    scale_y_continuous(limits = ylims) +
    scale_colour_manual(values = colours) +
    scale_fill_manual(values = colours) +
    labs(colour = "Mixture component",
         fill = "Mixture component")

  if(plot_id != 6){
    plot_bs[[plot_id]] <- plot_bs[[plot_id]] + 
      theme(legend.position = "none",
            axis.title = element_blank(),
            axis.text = element_blank(),
            axis.ticks = element_blank())
    plot_bw[[plot_id]] <- plot_bw[[plot_id]] + 
      theme(legend.position = "none",
            axis.title = element_blank(),
            axis.text = element_blank(),
            axis.ticks = element_blank())
  } else if(plot_id == 6){
    plot_bs[[plot_id]] <- plot_bs[[plot_id]] + 
      theme(legend.position = "none",
            axis.title = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank())
    plot_bw[[plot_id]] <- plot_bw[[plot_id]] + 
      theme(legend.position = "none",
            axis.title = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank())
  }

  plot_ww[[plot_id]] <- ggplot(data, aes(x)) +
    stat_function(aes(colour = "Fluent"),
                  geom = "line", 
                  fun = plot_mix_comps,
                  linewidth = .25,
                  args = list(alpha, sigma, lam = 1 - theta_ww)) +
    stat_function(aes(colour = "Hesitant"),
                  geom = "line", 
                  fun = plot_mix_comps, 
                  linewidth = .25,
                  args = list(alpha2_ww, sigmap_ww, lam = theta_ww)) +
    stat_function(aes(fill = "Fluent"),
                  geom = "area", 
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha, sigma, lam = 1 - theta_ww)) +
    stat_function(aes(fill = "Hesitant"),
                  geom = "area", 
                  fun = plot_mix_comps, 
                  alpha = .25, 
                  args = list(alpha2_ww, sigmap_ww, lam = theta_ww)) +
    scale_x_continuous(labels = ~exp(.) / 1000, 
                       breaks = xbreaks, 
                       limits = xlims) +
    scale_y_continuous(limits = ylims) +
    scale_colour_manual(values = colours) +
    scale_fill_manual(values = colours) +
    facet_grid(dataname~.) +
    labs(colour = "Mixture component",
         fill = "Mixture component")

   if(plot_id != 6){
    plot_ww[[plot_id]] <- plot_ww[[plot_id]] + 
      theme(legend.position = "top",
            axis.title = element_blank(),
            axis.text = element_blank(),
            axis.ticks = element_blank())
  } else{
    plot_ww[[plot_id]] <- plot_ww[[plot_id]] + 
      theme(legend.position = "top",
            axis.title = element_blank(), 
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank())
  }
}

bs_lab <- ggplot(mapping = aes(x = 0, y = 0, label = subtitle_bs)) + 
  geom_text() +
  theme_void()

bw_lab <- ggplot(mapping = aes(x = 0, y = 0, label = subtitle_bw)) + 
  geom_text() +
  theme_void()

ww_lab <- ggplot(mapping = aes(x = 0, y = 0, label = subtitle_ww)) + 
  geom_text() +
  theme_void()

# combine plots
plots_fin <- wrap_plots(
  wrap_plots(bs_lab, wrap_plots(plot_bs, ncol = 1), heights = c(.025, .975)),
  wrap_plots(bw_lab, wrap_plots(plot_bw, ncol = 1), heights = c(.025, .95)),
  wrap_plots(ww_lab, wrap_plots(plot_ww, ncol = 1), heights = c(.025, .975)), guides =  "collect")

# add axes titles
grid.arrange(patchworkGrob(plots_fin), 
                          left = "Density",
                          bottom = "Interkey interval in secs")
#ggsave("mixdists.png", plots_fin, width = 12, height = 6)

```

## Location estimates by model parameter

```{r message=FALSE}
files <- list.files(str_c("../../stanout/", 
                          c("lift", "spl2", "plantra", "cato", "c2l1", "gunnexp2")), 
                    pattern = "mogbetaco.+.csv", 
                    full.names = T)

ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode_factor(., 
                "c2l1" = "C2L1",
                "cato" = "CATO",# (non-dyslexic, unmasked)",
                "gunnexp2" = "GE2",# (unmasked)",
                "lift" = "LIFT",
                "plantra" = "PLanTra",
                "spl2" = "SPL2", # (L1)", 
#                "spl2 (shift + C)" = "SPL2 (L1; shift + C)", 
                .ordered = TRUE))) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES" & !is.na(lang)),
         !(str_detect(data, "CATO") & group == "dyslexic" & !is.na(group)),
         !(str_detect(data, "CATO") & task == "masked" & !is.na(task)),
         !(str_detect(data, "GE2") & xn == "masked" & !is.na(xn)))

ps_beta <- filter(ps, param == "beta") %>% 
  rename(beta = value) %>%
  mutate(idx = 1:n(), .by = data) %>% 
  select(idx, beta, data)

ps <- filter(ps, param != "beta")  %>% 
  pivot_wider(names_from = param, 
              values_from = value) %>% 
  unnest(cols = beta2:theta) %>%
  mutate(idx = 1:n(), .by = c(data, location, task, xn, genre, topic, lang)) %>% 
#  summarise(across(idx, list(max = max)), .by = c(location, data, task))
  left_join(ps_beta, by = c("data", "idx")) %>%
  mutate(across(c(beta, beta2), ~exp(.)),
         delta = beta2 - beta,
         across(location, ~str_replace(., " ", "\n"))) %>% 
  pivot_longer(c(beta, beta2, delta, prob)) %>% 
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper),
                   .names = "{.fn}"), 
            .by = c(data, location, name)) %>% 
  filter(name %in% c("beta", "beta2", "delta", "prob"))
```

```{r crossstudypost2, fig.cap="", fig.height=4.5, fig.width=10}

posd <- position_dodge(.5)
dotsize <- 2.75
grouplabel <- "Dataset:"
shapes <- c(1, 4, 6, 7, 8, 9, 11)
width <- 0
linewidth <- .35
nrow <- 1
beta <- filter(ps, name == "beta", 
               location == "before\nword") %>%
  mutate(name = beta_label,
         location = "overall") %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  facet_grid(~name, scales = "free_x") +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.title = element_blank()) +
  labs(colour = grouplabel,
       shape = grouplabel) +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label,
         across(c(mean, lower, upper), ~./1000)) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = scales::comma) +
  labs(colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title = element_blank()) +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title = element_blank(),
        legend.position = "top") +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

plot <- beta + delta + theta + plot_layout(guides = "collect",
                                           widths = c(.5, 1, 1))

grid.arrange(patchworkGrob(plot), 
             left = "Posterior estimate with 95% PIs",
             bottom = "Key location in text")

```

## Location estimates by model parameter

```{r crossstudypost3, fig.cap="", fig.height=4.5, fig.width=10}
beta <- beta +
  geom_line(linewidth = linewidth, 
            alpha = .75, 
            position = posd) +
  geom_errorbar(width = width, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd)

delta <- delta + 
  geom_line(linewidth = linewidth, alpha = .75, position = posd) +
  geom_errorbar(width = width, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd)

theta <- theta + 
  geom_line(linewidth = linewidth, alpha = .75, position = posd) +
  geom_errorbar(width = width, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd)

plot <- beta + delta + theta + plot_layout(guides = "collect",
                                           widths = c(.5, 1, 1))

grid.arrange(patchworkGrob(plot), 
             left = "Posterior estimate with 95% PIs",
             bottom = "Key location in text")

```

## Location estimates by model parameter

```{r loceffectconbf, results='asis'}

files <- list.files(str_c("../../stanout/", 
                          c("lift", "spl2", "gunnexp2", "plantra", "cato", "c2l1")), pattern = "mogbetacon.+.csv", full.names = T) # "spl2_shift",

ps_loc_diffs <- purrr::map_dfr(files, ~read_csv(.x) %>% 
                                 mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode(., 
                              "cato" = "CATO",
                              "spl2" = "SPL2", 
                              "gunnexp2" = "GE2",
                              #                "spl2_shift" = "*SPL2* (L1; shift + C)", 
                              "plantra" = "PLanTra",
                              "lift" = "LIFT",
                              "c2l1" = "C2L1"))) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked"),
         !(str_detect(data, "GE2") & xn == "masked"),
         param %in% c("delta", "theta")) %>%
  select(param, location, data, value) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, .))) %>% 
  pivot_wider(names_from = location, values_from = value) %>%
  unnest(-c(data, param)) %>% 
  mutate(diff.1 = `before sentence` - `before word`,
         diff.2 = `before word` - `within word`) %>% 
  summarise(across(starts_with("diff"), 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper, 
                        BF = BF)),
            .by = c(data, param))

tmp <- ps_loc_diffs %>% 
  pivot_longer(starts_with("diff"), names_to = c("diffid", ".value"), names_sep = "_") %>% 
  mutate(across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2))))) %>% 
  select(-mean, -lower, -upper) %>% 
  pivot_wider(names_from = c(diffid, param), values_from = BF) %>% 
  select(data, contains("delta"), contains("theta")) %>% 
  arrange(data)

tmp %>% 
  kable(col.names = c("Dataset", rep(c("before sentence vs word", "before vs within word"), 2)),
        caption = "Values indicate BFs in support of the alternative hypothesis over the null hypothesis.", align = c("l", rep("r", 4))) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 14) %>% 
  add_header_above(c(" " = 1,
                     "Hesitation slowdown" = 2,
                     "Hesitation probability" = 2))

```