### 2024-06-13 slides/sections/modelfit.Rmd

## Model fit to data

```{r getdensity, eval = F}
theme_set(ggthemes::theme_few(base_size = 8) +
              theme(strip.background = element_blank(),
                    legend.position = "top",
                    panel.grid = element_blank(),
                    panel.background = element_rect(fill = "transparent"), # bg of the panel
                    plot.background = element_rect(fill = "transparent", color = NA),
                    axis.text.y = element_blank(),
                    axis.ticks.y = element_blank(),
                    axis.title = element_blank()))

file <- list.files("../../stanout", pattern = "sims", full.names = T, recursive = T)
ds <- str_remove_all(file, "\\.|\\/|stanout|all_sims.csv") %>% 
    str_to_upper() %>% 
    str_replace(., "PLANTRA", "PLanTra") %>%
    str_replace(., "GUNNEXP2", "GE2")

# remove the shift model
ds <- ds[!str_detect(ds, "SHIFT")]

# get all density plots
plots <- map(1:length(ds), ~get_pdens_plot(file[.], max_iki = 5000))

# remove strips after first row
plots[2:length(ds)] <- map(2:length(ds), ~plots[[.]] + theme(strip.text = element_blank()))

# create labels for datasets
labs <- map(ds, ~ggplot(mapping = aes(x = 0, y = 0, label = .x)) + 
  geom_text(angle = 270, size = 2.5) +
  theme_void())

# add names for data sets on the right
plots_labs <- map(1:length(ds), ~plots[[.]] + labs[[.]] + plot_layout(widths = c(.95, .05)))

# combine legends
plots_ledg <- wrap_plots(plots_labs, guides = "collect", ncol = 1)

# add axe titles
plots_fin <- grid.arrange(patchworkGrob(plots_ledg), 
                            left = "Density", bottom = "Interkey interval in secs")

ggsave("../misc/gfx/fit.png", plot = plots_fin, width = 6.25, height = 8.5)

```

```{r fitplot, fig.cap="", eval = T}
include_graphics("../misc/gfx/fit.png")
```