### 2024-06-13 slides/sections/modeloverview.Rmd

## Model overview

```{r models, results = 'asis'}
tibble(Models = str_c("M",1:4),
       Type = c("LMM", "LMM", "LMM", "MoG"),
       Description = c("Single distribution Gaussian",
                       "Single distribution log-Gaussian",
                       "Single distribution log-Gaussian with different variance components per text location",
                       "Two-distributions mixture of log-Gaussians.")) %>% 
  select(Models, Description) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 18) %>% 
  group_rows("Serial", 1, 3) %>% 
  group_rows("Parallel", 4, 4) %>% 
  row_spec(0:4, extra_css = "line-height: 2.25;")  %>% 
  column_spec(1, width = "3cm")

```