### 2024-06-13 slides/sections/mog_app.Rmd

---

::: {style="margin-top:-20px; width: 96%;"}
```{r contpred}
inputPanel(
  sliderInput("beta_0", "Fluent typing", min = 100, max = 500, value = 150, ticks = FALSE),
  sliderInput("delta", "Hesitation slowdown", min = 50, max = 300, value = 100, ticks = FALSE),
  sliderInput("theta", "Hesitation probability", min = .001, max = .999, value = .3, ticks = FALSE),  
  switchInput("params", 
              size = "mini", 
              label = "Show means",
              onLabel = "on", 
              offLabel = "off", 
              labelWidth = "50%",
              width = "50%")
)

renderPlot({
  theme_set(theme_bw(base_size = 16) +
              theme(legend.position = "bottom",
                    legend.justification = "right",
                    legend.key.width = unit(.75, "cm"),
                    panel.grid = element_blank(),
                    axis.title = element_text(hjust = 0)))

  plot_mix_comps <- function(x, mu, sigma, lam) lam * dlnorm(x, mu, sigma)
  size <- .25
  alpha <- .25

  beta_0 <- log(200)
  beta_0 <- log(input$beta_0)
  delta <- 100
  delta <- input$delta
  theta <- .01
  theta <- input$theta
  beta_1 <- log(exp(beta_0) + delta)
  sigma <- .3
  sigma_diff <- .01
  sigmae <- sigma
  sigmae_p <- sigma + sigma_diff
  data <- data.frame(x = c(0, 1500))

  # Formula shown in plot
  formula <- list(~'iki'[i]~"~"~theta~paste("\u00B7")~  "logN("*beta~+~delta*","~sigma[2]^2*")",
                  ~+~"(1" - theta*")"~paste("\u00B7")~"logN("*beta*","~sigma[1]^2*")")

  p2 <- ggplot() +
    annotate("text", 
             x = 1.5, 
             y = seq(1, .25, 
                     length = length(formula)),
             label = formula, 
             parse = TRUE, 
             size = 6) +
    geom_rect(aes(xmin = -Inf, 
                  xmax = Inf, 
                  ymin = -Inf, 
                  ymax = Inf), 
              fill = "blue", 
              alpha = 0.05, 
              color = "black") +
    theme_void() +
    coord_cartesian(xlim = c(0, 3), ylim = c(-2.25, 2.4))

  if(input$params){
    beta_value <- round(exp(beta_0),0)
    theta_value <- dezero(theta, 2)
    delta_value <- round(delta, 2)
    beta_label <- substitute(beta==beta0, list(
      beta0 = beta_value))
    theta_label <- substitute(theta==theta_value, list(
      theta_value = theta_value))
    delta_label <- substitute(delta==delta_value, list(
      delta_value = delta_value))

    p2 <- p2 + 
      annotate("text",
               x = 2.25,
               y = -1,
               label = deparse(beta_label), 
               parse = TRUE, 
               size = 6) +
      annotate("text",
               x = 2.25,
               y = -1.5,
               label = deparse(theta_label), 
               parse = TRUE, 
               size = 6) +
      annotate("text",
               x = 2.25,
               y = -2,
               label = deparse(delta_label), 
               parse = TRUE, 
               size = 6)

  }

  p1 <- ggplot(data, aes(x)) +
    stat_function(geom = "line", fun = plot_mix_comps,
                  size = size,
                  args = list(beta_0, sigmae, lam = 1 - theta)) +
    stat_function(geom = "line", fun = plot_mix_comps,
                  size = size,
                  args = list(beta_1, sigmae_p, lam = theta)) +
    geom_area(stat = "function", 
              fun = plot_mix_comps, 
              show.legend = T, 
              alpha = alpha, 
              size = size, 
              aes(fill = "Fluent typing"), 
              args = list(beta_0, sigmae, lam = 1 - theta)) +
    geom_area(stat = "function", 
              fun = plot_mix_comps, 
              show.legend = T, 
              alpha = alpha, 
              size = size,
              aes(fill = 'Hesitation ("Pause")'), 
              args = list(beta_1, sigmae_p, lam = theta)) +
    geom_segment(x = exp(beta_0), 
                 y = 0, 
                 xend = exp(beta_0), 
                 yend = Inf, 
                 colour = "darkred", 
                 size = 1,
                 alpha = alpha) +
    geom_segment(x = exp(beta_1), 
                 y = 0, 
                 xend = exp(beta_1), 
                 yend = Inf, 
                 colour = "darkred", 
                 size = 1,
                 alpha = alpha) +
    geom_segment(x = exp(beta_0),
                 y = 0.0065, 
                 xend = exp(beta_1), 
                 yend = 0.0065, 
                 colour = "blue", 
                 alpha = alpha,
                 size = 1,
                 arrow = arrow(length = unit(0.3, "cm"), 
                               ends = "both", 
                               type = "closed")) +
    scale_fill_manual(breaks = c("Fluent typing", 'Hesitation ("Pause")'),
                      values = c("#000000","#009E73")) +
    scale_x_continuous(labels = scales::comma, breaks = seq(0, 2500, 200)) +
    coord_cartesian(ylim = c(0, .007), 
                    xlim = c(0, 800)) +
    labs(title = "",
         fill = "Mixture components",
         x = "Inter-keystroke interval [in msecs]",
         y = "Density")

  p1 + p2 + 
    plot_layout(widths = c(3, 1.25)) + 
    plot_layout(guides = 'collect')

})

```
:::