### 2024-06-13 slides/sections/sampledata.Rmd

## Some sample data

Essay-writing task from ProWrite project [e.g. @dux2022automating] without deletions and spaces.

```{r eval = F}
read_csv('tutorial/data/process_data_by_event.csv') %>% 
  filter(location != "post-paragraph", 
         !is.na(location),
         next_event_type == "production",
         event_type == "production",
         event_duration > 50,
         event_duration < 30000,
         session_no == 1) %>%
  drop_na(text) %>%
  mutate(across(text, str_to_lower)) %>% 
  filter(text %in% letters) %>% 
  mutate(.by = c(token, location), 
         idx = sample(n())) %>% 
  filter(idx <= 50) %>% 
  select(participant, iki = event_duration, location, pretext, text) %>% 
  mutate(across(location, ~recode(., 
                                  "within-word" = "within word",
                                  "post-word" = "before word",
                                  "post-sentence" = "before sentence"))) %>% 
  filter(participant %in% sample(unique(participant), 10)) %>% 
  mutate(recent_text = substr(pretext, 
                              start = nchar(pretext) - 40,
                              stop = nchar(pretext))) %>% 
  select(-pretext) %>% 
  rename(new_letter = text) %>% 
  relocate(participant, location, recent_text, new_letter, iki) %>% 
  write_csv("slides/data/sampledata.csv")
```

```{r}
data <- read_csv("slides/data/sampledata.csv")
slice(data, sample(nrow(data), size = 10))
```

## How are these data distributed?

```{r fig.width=10, fig.height=6}
ggplot(data, aes(x = iki, colour = location, fill = location)) +
  geom_density(alpha = .25) +
  scale_colour_colorblind() +
  scale_fill_colorblind() +
  scale_x_log10() 
```