### 2024-06-13 slides/sections/simulation.Rmd

```{r loossim, results='asis'}
table <- read_csv("../../mixture-model-sim/stanout/modelcomparison.csv") %>% 
  select(1:6) %>% 
  mutate(
#    elpd_se_ratio = abs(elpd_diff / se_diff),
    across(ends_with("diff"), ~round(., 1)),
    across(ends_with("loo"), ~round(., 0)),
    across(where(is.numeric), ~format(., big.mark = ",")),
    across(everything(), ~as.character(.)),
    across(everything(), ~str_trim(.)),
    across(everything(), ~str_replace_all(., "NaN", "--")),
    across(elpd_diff, ~str_c(., " (", se_diff, ")")),
    across(elpd_loo, ~str_c(., " (", se_elpd_loo, ")")),
    across(elpd_diff, ~str_replace_all(., "0.0 \\(0.0\\)", "--")),
    across(c(data, model), ~recode(., mog = "Two-distribution mixture model",
                                      lm = "Single-distribution model"))) %>% 
  select(data, model, starts_with("elpd"))

# names(table) <- c("Data", "Model", "$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$")
# 
# table %>% 
#   select(-Data) %>% 
#   kable(align = c("l", "r", "r")) %>% 
#   kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
#                 full_width = T, 
#                 position = "right",
#                 font_size = 16) %>% 
#   group_rows("Two-distributions data", 1, 2) %>% 
#   group_rows("Single-distribution data", 3, 4) 
mixturedata <- table %>% filter(str_detect(data, "Two"), elpd_diff != "--") %>% pull(elpd_diff)
singledata <- table %>% filter(str_detect(data, "Single"), elpd_diff != "--") %>% pull(elpd_diff)
```

## Simulation {.smaller}

<div style="float: left; width: 43%;">
\
\

- Simulate data from a single log-normal distribution
- Simulate data from a mixture of to log-normal distribution
- Data simulated with mixture process: 
  - Advantage for mixture model over single process model: $\Delta\widehat{elpd} =$ `r mixturedata`
- Data simulated with single process: 
  - Negligible difference between models: $\Delta\widehat{elpd} =$ `r singledata`

</div>

<div style="float: right; width: 50%;">
```{r simdata, fig.cap= "", fig.width=4, fig.height=4, out.width = "90%"}
files <- list.files("../../mixture-model-sim/data", full.names = T)

map_dfr(files, ~read_csv(.) %>% 
        mutate(Dataset = str_remove_all(.x, "\\../|/|mixture-model-sim/data|/|\\.|csv"))) %>%
  mutate(across(Dataset, ~recode(., mogdata = "two distributions", lmdata = "single distribution"))) %>% 
  ggplot(aes(x = value, colour = Dataset, fill = Dataset)) +
  geom_density(alpha = .25, 
               linewidth = .25) +
  scale_colour_colorblind() + 
  scale_fill_colorblind() +
  scale_x_log10(labels = ~format(., big.mark = ",")) +
  theme(text = element_text(size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  labs(colour = "Data", 
       fill = "Data", 
       x = "y")
```
</div>

```{r eval = F}
## Parameter estimates

mog1 <- read_csv("../../mixture-model-sim/stanout/mog_mogdata.csv") %>% 
  mutate(data = "mog")
mog2 <- read_csv("../../mixture-model-sim/stanout/mog_lmdata.csv") %>% 
  mutate(data = "lm")
lm1 <- read_csv("../../mixture-model-sim/stanout/lm_lmdata.csv") %>% 
  mutate(data = "lm")
lm2 <- read_csv("../../mixture-model-sim/stanout/lm_mogdata.csv") %>% 
  mutate(data = "mog")

mog_sim <- bind_rows(mog1, mog2) %>% 
  summarise(across(value, list(est = mean,
                               lo = lower,
                               up = upper),
                   .names = "{.fn}"),
            .by = c(name, data)) %>% 
  mutate(across(where(is.numeric), ~round(., 2)),
         across(where(is.numeric), ~ifelse(str_detect(name, "prob"), dezero(., 2), .)),
         across(est, ~str_c(., " [", lo, ", ", up, "]")),
         across(name, ~str_c("\\", .)),
         across(name, ~str_replace(., "a_e", "a^2_1")),
         across(name, ~str_replace(., "ap_e", "a^2_2"))) %>%
  select(-lo, -up) %>% 
  pivot_wider(names_from = data, values_from = est) %>% 
  mutate(true = c(5, 1, .35, .25, .5),
         across(true, ~ifelse(str_detect(name, "prob"), dezero(., 2), .)),
         across(name, ~str_replace(., "prob", "theta"))) %>% 
  relocate(name, true) %>% 
  unite("name", name:true, sep = " = ")

lm_sim <- bind_rows(lm1, lm2) %>% 
  summarise(across(value, list(est = mean,
                               lo = lower,
                               up = upper),
                   .names = "{.fn}"),
            .by = c(name, data)) %>% 
  mutate(across(where(is.numeric), ~round(., 2)),
         across(est, ~str_c(., " [", lo, ", ", up, "]")),
         across(name, ~str_c("\\", .)),
         across(name, ~str_replace(., "a_e", "a^2_1")),
         across(name, ~str_replace(., "ap_e", "a^2_2"))) %>% 
  select(-lo, -up) %>%
  pivot_wider(names_from = data, values_from = est) %>% 
  mutate(true = as.character(c(5, .25))) %>% 
  relocate(name, true) %>% 
  unite("name", name:true, sep = " = ")

```

```{r simparam, results='asis', eval = F}
table <- bind_rows(mog_sim %>% mutate(model = "mog"),
                   lm_sim %>% mutate(model = "lm")) %>% 
  relocate(model) %>% 
  mutate(across(name, ~str_c("$", ., "$")))

names(table) <- c("Model", "Parameter values used in simulation", "Two-distributions model", "Single-distributions model")

table %>% 
  select(-Model) %>% 
  kable(align = c("l", "r", "r")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 16) %>% 
  group_rows("Two-distributions data", 1, 5) %>% 
  group_rows("Single-distribution data", 6, 7) %>% 
  add_header_above(c(" " = 1,
                     "Estimate with 95\\% PI" = 2))

```