### 2024-06-13 slides/sig-paris-2024-keynote/main.Rmd

---
title: '<span style="font-size: 50px; font-face: bold">Modelling cascaded planning in written production as Bayesian hierarchical mixture process</span>'
author: '<span style="font-size: 30px; font-face: bold">Jens Roeser (Nottingham Trent University)</span>'
output: 
  ioslides_presentation:
    incremental: false
    transition: slower
    widescreen: true
    css: ../misc/slides.css
    self_contained: true
#    logo: ../misc/gfx/ntu.png
runtime: shiny
bibliography      : ["../../report/references.bib"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      eval = TRUE, 
                      message = FALSE, 
                      comment=NA,
                      fig.align = "right")
options("kableExtra.html.bsTable" = T, digits = 2)
options(pillar.print_min = 5, pillar.print_max = 10)
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(knitr)
library(kableExtra)
library(gridExtra)
library(patchwork)
library(ggthemes)
library(polspline)
source("../../scripts/functions.R")
source("../scripts/get_pdens_plot.R")

beta_label <- "Fluent interkey\nintervals (in msecs)"
delta_label <- "Slowdown for hesitant\ninterkey intervals (in secs)"
theta_label <- "Probability of hesitant\ninterkey intervals"

theme_set(theme_few(base_size = 12) +
            theme(strip.background = element_blank(),
                  legend.position = "top",
                  legend.justification = "right",
                  panel.grid = element_blank(),
                  panel.background = element_rect(fill = "transparent"), # bg of the panel
                  plot.background = element_rect(fill = "transparent", color = NA)))
```

## Keystroke interval data

```{r}
data_pseq_1 <- read_csv('tutorial/data/process_data_by_event.csv') %>% 
  filter(location != "post-paragraph", 
         !is.na(location),
         next_event_type == "production",
         event_type == "production",
         event_duration > 50,
         session_no == 1) %>% 
  filter(participant == unique(participant)[1], 
         pseq_id == 10, 
         session_no == 1) %>% 
  select(text, iki = event_duration, location) %>% 
  rowid_to_column("id")
```

```{r fig.width=10, fig.height=6}
d <- data_pseq_1 %>% 
  mutate(
    across(location, ~str_replace(., "post-", "before ")),
    across(location, ~str_replace(., "-", " ")),
    across(text, ~replace_na(., " ")))

ggplot(d, aes(y = iki, x = id, group = 1)) +
  geom_path() +
  geom_point(aes(colour = location, shape = location), size = 3.5) +
  theme_classic(base_size = 14) +
  labs(y = "Interkey intervals (in msecs)") +
  scale_x_continuous(breaks = d$id, labels = d$text) +
  scale_colour_colorblind() +
  scale_shape_manual(values = c(8, 17, 16)) +
  scale_y_log10() +
  labs(colour = "text locations",
       shape = "text locations") +
#  scale_y_continuous(limits = c(0, 1100)) +
  theme(axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "bottom")
```

# How are the processes that generate and sequences keystrokes mentally coordinated?

```{r intro, child = "../sections/introduction.Rmd"}
```

```{r intro, child = "../sections/evidence.Rmd"}
```

```{r intro, child = "../sections/sampledata.Rmd"}
```

## Usual practice

```{r echo = T}
data_with_pause <- data %>% 
  mutate(is_pause_500 = iki > 1000,
         is_pause_1k = iki > 1000,
         is_pause_2k = iki > 2000)
data_with_pause
```

```{r echo = T}
data_with_pause %>% 
  summarise(across(starts_with("is_pause"), list(avg = mean)), .by = location)
```

## Usual practice

```{r echo = T}
data_with_pause %>% 
  filter(!is_pause_500) %>% 
  summarise(iki_avg = mean(iki),
            iki_sd = sd(iki), 
            .by = location)
```

# How can we represent these models statistcally?

## Problems

- Pause thresholds are in principle arbitrary.
- Separation of distributions does not capture the process that generates keystroke data.

```{r data, child = "../sections/statisticalmodels.Rmd"}
```

```{r intro, child = "../sections/mog_app.Rmd"}
```

# How can we apply these formulae to data?

## Single process models

## Why Bayesian models?

```{r}

# mixture models map directly on parallel cascaded architecture of typing

# how did people usually deal with pauses -> log, threshold

# Why Bayesian framework?

# start with normal models and lm, lmer, brms

# show Bayesian inference

```

## Thank you!

This work was supported by UKRI ESRC under Grant No. ES/W011832/1: "Can you use it in a sentence?: Establishing how word-production
difficulties shape text formation." (Nottingham Trent University)

Data and analysis scripts written in R and Stan can be found on OSF https://osf.io/z65dw/ [@roeser2024osf].

For a tutorial on Bayesian mixture-model analysis in the context of keystroke data see
https://rpubs.com/jensroes/mixture-models-tutorial.

"Ideas cascading into keystrokes -- Modelling writing hesitations as Bayesian mixture process" June 27 Session D7 (D 321) 10.30am-12.00pm

Roeser, J., De Maeyer, S., Leijten, M., & Van Waes, L. (2024). Modelling typing disfluencies as finite mixture process. *Reading and Writing*, 37(2), 359-384.

## References {.smaller}

<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):after {
  content: '';
}
</style>