### 2024-10-26 report/appendix/l2_constr.Rmd

# L2 effect (SPL2)

For the SPL2 data we calculated the L2 effect (i.e. the effect of writing in L2 or L1). The results can be found in Table \ref{tab:l2effect}. The results show longer hesitation duration and more hesitations across all transition locations when writing in L2. Only within-word transitions showed negligible evidence for more or longer hesitations.

```{r l2effect, results = 'asis'}
cellmeans <- ps_spl2_both %>% 
  mutate(across(name, ~recode(., beta2 = "delta"))) %>% 
  separate(group, into = c("group", "lang"), sep = "; ") %>%
  select(-ds, -group) %>%
  mutate(across(lang, ~str_remove(., "Language: ")),
         across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, data, lang), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = data, 
              values_from = est) %>% 
  mutate(across(c(location, lang), ~ifelse(param == "beta", "overall", .))) %>% 
  unique()

# cell means and data set difference
cellmeans <- cellmeans %>% 
  select(-`SPL2 (_^[shift] + C)`) %>% 
  pivot_wider(names_from = lang, values_from = `SPL2 (_^[shift])`)

file1 <- list.files(path = "../../stanout/spl2", 
                    pattern = "mogbetaco.+.csv$",
                    full.names = T)

tmp <- map_dfr(file1, read_csv) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  mutate(across(lang, recode, "EN" = "L1", "ES" = "L2")) %>% 
  pivot_wider(names_from = lang, values_from = value) %>% 
  unnest(c(L1, L2)) %>% 
  mutate(diff = L2 - L1) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, recode, theta = "prob")) %>% 
  left_join(cellmeans, by = join_by(location, param)) %>% 
  arrange(param, location) %>% 
  select(param, `Transition location` = location, L1, L2,
         "Difference" = mean, BF)

write_csv(tmp, "../tables/l2_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for language effect. Cellmeans are shown for the hesitation duration and for the probability of hesitant interkey intervals for texts produced in the writer's L1 and L2. Language differences are shown for the hesitation duration (log scale) and the hesitation probability (logit scale); 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          font_size = "small",
          stub_indents = list("Hesitation duration" = 1:3,
                              "Hesitation probability" = 4:6) )

```