### 2024-10-26 report/appendix/masking_constr.Rmd

# Masking effect (CATO, GE2)

Studies associated with the CATO and GE2 datasets investigated to what extent hiding (masking) previously written text from the reader affects keystroke behaviour. Mixture model results for the masking effect are shown in Table \ref{tab:maskingeffect}. There is some evidence that when the text was masked dylexic writers hesitated longer before starting to type sentences. Evidence for all other comparisons was negligible.

\blandscape
```{r maskingeffect, results = 'asis'}
ps_cato_2 <- ps_cato %>% 
  separate(group, into = c("mask","group"), sep = "; ") %>% 
  mutate(across(c(mask, group), ~str_remove(., "Task: |Group: ")))

cellmeans_mask <- ps_gunn %>% 
  rename(mask = group) %>%
  mutate(group = "non dyslexic") %>% 
  bind_rows(ps_cato_2) %>%
  mutate(across(name, ~recode(., beta2 = "delta"))) %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, group, mask, ds), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = mask, 
              values_from = est)

file <- list.files(path = "../../stanout/gunnexp2", 
                   pattern = "mogbetaco.*.csv$",
                   full.names = T)

d_gunn <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value)),
         ds = "GE2",
         across(param, ~recode(., "prob" = "theta")),
         group = "non dyslexic") %>% 
  rename(mask = xn)

# get posterior
file <- list.files(path = "../../stanout/cato", 
                   pattern = "mogbetaco.+.csv$",
                   full.names = T)

d_cato <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "prob")) %>% 
  mutate(ds = "cato") %>% 
  rename(mask = task)

tmp <- bind_rows(d_gunn, d_cato) %>% 
  pivot_wider(names_from = mask, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param, group, ds)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob")),
         across(ds, ~str_to_upper(.))) %>% 
  left_join(cellmeans_mask, by = join_by(location, param, group, ds)) %>% 
  arrange(param, location, ds, group) %>% 
  select(param, `Transition location` = location,
         `Dataset` = ds, Group = group, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "../tables/masking_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "l", "l", "r", "r", "r", "r"), 
          caption = "Mixture-model estimates for masking effect. Cellmeans are shown for the masked and unmasked writing task for hesitation duration (in msecs) and hesitation probability. The masking effect is shown on the log scale for durations and on the logit scale for probabilities; 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Hesitation duration" = 1:9,
            "Hesitation probability" = 10:18) )

```
\elandscape