### 2024-10-26 report/appendix/prepost_constr.Rmd

# Pre-post test (PLanTra)

The pre-post test effect for the PLanTra dataset is reported in Table \ref{tab:retesteffect}. Evidence for all comparisons was negligible.

```{r retesteffect, results = 'asis'}
cellmeans <- ps_plantra %>% 
  mutate(across(name, ~recode(., beta2 = "delta"))) %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(location, group, ds), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper, -ds) %>% 
  pivot_wider(names_from = group, values_from = est)

file <- list.files(path = "../../stanout/plantra", 
                   pattern = "mogbetaco.*.csv$",
                   full.names = T)

tmp <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  pivot_wider(names_from = task, values_from = value) %>% 
  unnest(c(`post-test`,`pre-test`)) %>% 
  mutate(diff = `pre-test` - `post-test`) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob"))) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location,
         `Pre-test` = `pre-test`,
         `Post-test` = `post-test`, 
         "Difference" = mean, BF)

write_csv(tmp, "../tables/posttest_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for post-test effect. Cellmeans are shown for pre and post-test for the duration of hesitant interkey intervals in msecs and its probability. Differences are shown on the log scale for hesitation duration and on the logit scale for the hesitation probability; 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Hesitation duration" = 1:3,
            "Hesitation probability" = 4:6))

```