### 2024-10-26 slides/sections/discussion.Rmd

## Planning text unfolds in parallel to production!

- First robust evidence that written composition is largely a parallel process.
- Writers do not necessarily pause at larger linguistic locations but plan utterances in parallel to writing.
- Evidenced by 
  - stronger predictive performance of mixture models [@roeser2021modelling].
  - pauses are not generally more likely before sentences compare to words (but often longer).
- Pauses are consistently more likely at word-initial location compared to word-medial but (1) not always longer and (2) for sentence-initial interkey intervals pausing behaviour varies as a function of writing experience, languages (e.g. young / L2 writers, students), composition tasks.