### 2024-10-26 slides/sections/evidence.Rmd

## Serial view

- Pause duration before a writer starts a sentence is typically longer than before a mid-sentence word and these are longer than between mid-word key presses [e.g. @conijn2019understanding]. 
- Long interkey intervals are typically followed by production bursts [@hayes2012evidence; @kaufer1986composing].
- Planning message and syntax for a sentence adds to the time required to prepare the upcoming word, and the motor planning required to produce the first keystroke [@baaijen2012keystroke; @roeser2018advance].

## Parallel view

- Variations in interkey intervals are not sufficiently explained by text location.
- Writers often do not hesitate before sentence (or word) onset.
- In some cases sentences are not planned, planning was postponed until after sentence onset, or completed in parallel with previous output [@not10; @roeser2018advance].
- Sentence initial durations are often shorter than what one would expect if sentence-initial pauses reflect planning.
- Fluent output is maintained by a cascade of processes that run in parallel [@olive2014toward; @van1991handwriting].
  <!-- - @medimorec2017pauses: 71% shorter than 1 sec -->
  <!-- - @ronneberg2022process: 50% around 430 msecs and around 1.2 secs  -->
  <!-- - For comparison, these short pauses are less than mean written naming response latency for single objects [@torrance2018timed] and longer pauses were similar to response latencies for short sentences when describing arrays of images [@roeser2018advance].  -->