### 2024-10-26 slides/sections/implementation.Rmd

## Implementation and model comparison

<!-- - Bayesian inference [e.g. @lambert2018student] models implemented in Stan [@carpenter2016stan] using `rstan` [@rstan] to interface with $R$ -->
- Code based on @vasishth2017 [see @roeser2021modelling]
- Random effects for participants
- Leave-one-out cross-validation: sum of the *expected log predictive density* $\widehat{elpd}$ 
- *Difference between models* $\Delta\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical]
- Difference between models was summarised as

$\mid\frac{\Delta\widehat{elpd}}{\text{SE}_\text{diff}}\mid$,

i.e. the *z*-score of the change in predictive performance [@sivula2020uncertainty].