### 2024-10-26 slides/sections/modelcomparisons.Rmd

<!-- ## Model comparisons -->

```{r loadloos, warning=FALSE, message=FALSE}
files <- list.files(
  str_c("../../stanout/",
        c("lift", "spl2", "spl2_shift", "plantra", "cato", "c2l1", "gunnexp2")),
  pattern = "model.+.csv",
  full.names = T)

mc <- purrr::map_dfr(files, ~read_csv(.) %>%
                          mutate(dataset = .x)) %>%
  mutate(across(dataset, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .))) %>%
  filter(dataset != "spl2_shift",
         model != "M4 -- M5") %>%
  mutate(across(dataset, ~recode(.,
                            "cato" = "CATO",
                            "spl2" = "SPL2",
                            #                "spl2_shift" = "SPL2 (shift + C)",
                            "plantra" = "PLanTra",
                            "lift" = "LIFT",
                            "c2l1" = "C2L1",
                            "gunnexp2" = "GE2")),
         across(model, ~case_when(str_detect(., "M2$") ~ "Single process (M1 vs M2)",
                               str_detect(., "M3$") ~ "Single process (unequal var.; M2 vs M3)",
                               str_detect(., "M4$") ~ "Mixture process (M3 vs M4)",
                             )),
         across(model, ~factor(., levels = unique(model), ordered = T)))
```

<!-- ```{r modelcomp, fig.cap="Shown is the difference in predictive performance of models."} -->
<!-- plot <- mc %>%  -->
<!--   mutate(value = elpd_diff, -->
<!--          se = se_diff, -->
<!--          lower = value - 2 * se, -->
<!--          upper = value + 2 * se) %>%  -->
<!--   select(model, dataset, value, lower, upper) %>%  -->
<!--   ggplot(aes(x = reorder(dataset, value), -->
<!--              y = value,  -->
<!--              ymin  = lower,  -->
<!--              ymax = upper, -->
<!--              colour = model)) + -->
<!--   facet_wrap(~model, scales = "free_x", nrow = 2) + -->
<!--   coord_flip() + -->
<!--   theme_bw() + -->
<!--   scale_y_continuous(labels = scales::comma) + -->
<!--   ggthemes::scale_color_colorblind() + -->
<!--   labs(y = bquote(Delta*widehat(elpd)~"with 2 \u00D7 SE")) + -->
<!--   theme(axis.title.y = element_blank(), -->
<!--         axis.ticks.y = element_blank()); plot -->
<!-- ``` -->
<!-- ## Model comparisons -->

<!-- ```{r modelcomp2, fig.cap="Shown is the difference in predictive performance of models."} -->
<!-- plot + geom_pointrange(show.legend = F) -->
<!-- ``` -->

## Model comparisons

```{r loos, results='asis'}
mc %>% 
  select(dataset, Comparison = model, elpd_diff_ratio) %>%
  pivot_wider(names_from = Comparison, values_from = elpd_diff_ratio) %>%
  mutate(across(-dataset, ~" ")) %>% 
  rename(`Data set` = dataset) %>% 
  select(1, 4, 3, 2) %>% 
  kable(align = c("l", rep("r", 3)),
        caption = "Values are the absolute ratio of the difference in predictive performance measured as $\\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical] and its standard error $\\mid\\frac{\\Delta\\widehat{elpd}}{\\text{SE}}\\mid$ which corresponds to the $z$-score of the change in predictive performance [@sivula2020uncertainty].") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 14) %>% 
  row_spec(0:6, extra_css = "line-height: 2.25;")  
```

## Model comparisons

```{r loos2, results='asis'}
mc %>% 
  select(dataset, Comparison = model, elpd_diff_ratio) %>%
  pivot_wider(names_from = Comparison, values_from = elpd_diff_ratio) %>%
  rename(`Data set` = dataset) %>% 
  select(1, 4, 3, 2) %>% 
  kable(align = c("l", rep("r", 3)),
        caption = "Values are the absolute ratio of the difference in predictive performance measured as $\\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical] and its standard error $\\mid\\frac{\\Delta\\widehat{elpd}}{\\text{SE}}\\mid$ which corresponds to the $z$-score of the change in predictive performance [@sivula2020uncertainty].") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = T, 
                position = "right",
                font_size = 14) %>% 
  row_spec(0:6, extra_css = "line-height: 2.25;")  
```

```{r eval = F}
mc %>% 
  select(dataset, model, elpd = elpd_diff_ratio) %>% 
  ggplot(aes(y = elpd, x = model, colour = dataset, group = dataset)) +
  geom_point() +
  geom_line()
```