### 2024-10-26 slides/sections/mog_app.Rmd

# Mixture model parameters

---

::: {style="margin-top:-20px; width: 96%;"}
```{r contpred}
inputPanel(
  sliderInput("beta_0", "Fluent typing", min = 50, max = 500, value = 300, step = 10, ticks = FALSE),
  sliderInput("delta", "Hesitation slowdown", min = 50, max = 1500, value = 100, step = 50, ticks = FALSE),
  sliderInput("theta", "Hesitation probability", min = .001, max = .999, value = .3, step = .05, ticks = FALSE)  
  # switchInput("params", 
  #             size = "mini", 
  #             label = "Show means",
  #             onLabel = "on", 
  #             offLabel = "off", 
  #             labelWidth = "50%",
  #             width = "50%", 
  #             value = FALSE)
)

renderPlot({
  theme_set(theme_bw(base_size = 16) +
              theme(legend.position = "bottom",
                    legend.justification = "right",
                    legend.key.width = unit(.75, "cm"),
                    panel.grid = element_blank(),
                    axis.title = element_text(hjust = 0)))

  plot_mix_comps <- function(x, mu, sigma, lam) lam * dlnorm(x, mu, sigma)
  size <- .25
  alpha <- .25

  beta_0 <- log(200)
  beta_0 <- log(input$beta_0)
  delta <- 100
  delta <- input$delta
  theta <- .01
  theta <- input$theta
  beta_1 <- log(exp(beta_0) + delta)
  sigma <- .3
  sigma_diff <- .01
  sigmae <- sigma
  sigmae_p <- sigma + sigma_diff
  data <- data.frame(x = c(0, 2500))

  # Formula shown in plot
  # formula <- list(~'iki'[i]~"~"~theta~paste("\u00B7")~  "logN("*beta~+~delta*","~sigma[2]^2*")",
  #                 ~+~"(1" - theta*")"~paste("\u00B7")~"logN("*beta*","~sigma[1]^2*")")
  # 
  # p2 <- ggplot() +
  #   # annotate("text", 
  #   #          x = 1.5, 
  #   #          y = seq(1, .25, 
  #   #                  length = length(formula)),
  #   #          label = formula, 
  #   #          parse = TRUE, 
  #   #          size = 7) +
  #   geom_rect(aes(xmin = -Inf, 
  #                 xmax = Inf, 
  #                 ymin = -Inf, 
  #                 ymax = Inf), 
  #             fill = "blue", 
  #             alpha = 0.05, 
  #             color = "black") +
  #   theme_void() +
  #   coord_cartesian(xlim = c(0, 3), ylim = c(-2.25, 2.4))

  # if(input$params){
  #   beta_value <- round(exp(beta_0),0)
  #   theta_value <- dezero(theta, 2)
  #   delta_value <- round(delta, 2)
  #   beta_label <- substitute(beta==beta0, list(
  #     beta0 = beta_value))
  #   theta_label <- substitute(theta==theta_value, list(
  #     theta_value = theta_value))
  #   delta_label <- substitute(delta==delta_value, list(
  #     delta_value = delta_value))
  #   
  #   
  #   p2 <- p2 + 
  #     annotate("text",
  #              x = 2.25,
  #              y = -1,
  #              label = deparse(beta_label), 
  #              parse = TRUE, 
  #              size = 6) +
  #     annotate("text",
  #              x = 2.25,
  #              y = -1.5,
  #              label = deparse(theta_label), 
  #              parse = TRUE, 
  #              size = 6) +
  #     annotate("text",
  #              x = 2.25,
  #              y = -2,
  #              label = deparse(delta_label), 
  #              parse = TRUE, 
  #              size = 6)
  #   
  # }

  ggplot(data, aes(x)) +
    stat_function(geom = "line", fun = plot_mix_comps,
                  size = size,
                  args = list(beta_0, sigmae, lam = 1 - theta)) +
    stat_function(geom = "line", fun = plot_mix_comps,
                  size = size,
                  args = list(beta_1, sigmae_p, lam = theta)) +
    geom_area(stat = "function", 
              fun = plot_mix_comps, 
              show.legend = T, 
              alpha = alpha, 
              size = size, 
              aes(fill = "Fluent typing"), 
              args = list(beta_0, sigmae, lam = 1 - theta)) +
    geom_area(stat = "function", 
              fun = plot_mix_comps, 
              show.legend = T, 
              alpha = alpha, 
              size = size,
              aes(fill = 'Hesitation ("Pause")'), 
              args = list(beta_1, sigmae_p, lam = theta)) +
    geom_segment(x = exp(beta_0), 
                 y = 0, 
                 xend = exp(beta_0), 
                 yend = Inf, 
                 colour = "darkred", 
                 size = 1,
                 alpha = alpha) +
    geom_segment(x = exp(beta_1), 
                 y = 0, 
                 xend = exp(beta_1), 
                 yend = Inf, 
                 colour = "darkred", 
                 size = 1,
                 alpha = alpha) +
    geom_segment(x = exp(beta_0),
                 y = 0.0065, 
                 xend = exp(beta_1), 
                 yend = 0.0065, 
                 colour = "blue", 
                 alpha = alpha,
                 size = 1,
                 arrow = arrow(length = unit(0.3, "cm"), 
                               ends = "both", 
                               type = "closed")) +
    scale_fill_manual(breaks = c("Fluent typing", 'Hesitation ("Pause")'),
                      values = c("#000000","#009E73")) +
    scale_x_continuous(labels = scales::comma, breaks = seq(0, 2500, 200)) +
    coord_cartesian(ylim = c(0, .005), 
                    xlim = c(-10, 2100)) +
    labs(title = "",
         fill = "Mixture components",
         x = "Inter-keystroke interval [in msecs]",
         y = "Density")

#  p1 + p2 + 
#    plot_layout(widths = c(3, 1.25)) + 
#    plot_layout(guides = 'collect')

})

```
:::

## Parallel view: mixture of two processes {.smaller}

<div style="float: left; width: 40%;">

```{r fig.width=5}
# function for mixture models
plot_mix_comps <- function(x, mu, sigma, lam) lam * dnorm(x, mu, sigma)

# Parameters
data <- data.frame(x = c(50, 1500))
theta_conj <- .3
theta_simp <- .1
alpha <- 250
delta <- c(250, 500)
alpha2_w <- alpha + delta[1]
alpha2_s <- alpha + delta[2]
sigma <- 40
sigma_diff <- c(30, 60)
sigmae <- sigma
sigmae_p_w <- sigma + sigma_diff[1]
sigmae_p_s <- sigma + sigma_diff[2]

ggplot(data, aes(x)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = word_lbl),
                linewidth = linewidth,
                args = list(alpha, sigmae, lam = 1 - theta_simp)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = word_lbl),
                linewidth = linewidth,
                args = list(alpha2_w, sigmae_p_w, lam = theta_simp)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = word_lbl), alpha = .25, 
            args = list(alpha, sigmae, lam = 1 - theta_simp)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = word_lbl), alpha = .25, 
            args = list(alpha2_w, sigmae_p_w, lam = theta_simp)) +
  scale_color_manual(breaks = c(word_lbl, sent_lbl), values = ggthemes::colorblind_pal()(2)) +
  scale_fill_manual(breaks = c(word_lbl, sent_lbl), values = ggthemes::colorblind_pal()(2)) +
  scale_x_continuous(labels = scales::comma, 
                     limits = c(50, 1000), 
                     breaks = seq(0, 1500, 200)) +
  labs(x = "Interkey intervals [in ms]", 
       colour = "Text location",
       fill = "Text location") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

```
</div>

<div style="float: right; width: 50%;">
\
\
\
\

$$
\text{iki}_{i} \sim\text{ } \theta \times \text{log}\mathcal{N}(\beta + \delta, \sigma_{e'}^2) + \\
   (1 - \theta) \times \text{log}\mathcal{N}(\beta, \sigma_{e}^2)
$$

</div>

## Parallel view: mixture of two processes {.smaller}

<div style="float: left; width: 40%;">

```{r fig.width=5}
ggplot(data, aes(x)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = word_lbl),
                linewidth = linewidth,
                args = list(alpha, sigmae, lam = 1 - theta_simp)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = sent_lbl),
                linewidth = linewidth,
                args = list(alpha, sigmae, lam = 1 - theta_conj)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = word_lbl),
                linewidth = linewidth,
                args = list(alpha2_w, sigmae_p_w, lam = theta_simp)) +
  stat_function(geom = "line", fun = plot_mix_comps,
                aes(colour = sent_lbl),
                linewidth = linewidth,
                args = list(alpha2_s, sigmae_p_s, lam = theta_conj)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = word_lbl), alpha = .25, 
            args = list(alpha, sigmae, lam = 1 - theta_simp)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = sent_lbl), alpha = .15,
            args = list(alpha, sigmae, lam = 1 - theta_conj)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = word_lbl), alpha = .25, 
            args = list(alpha2_w, sigmae_p_w, lam = theta_simp)) +
  geom_area(stat = "function", fun = plot_mix_comps, aes(fill = sent_lbl), alpha = .15,
            args = list(alpha2_s, sigmae_p_s, lam = theta_conj)) +
  geom_curve(aes(x = alpha2_w, 
                 xend = alpha2_s, y = 0.0008, yend = 0.0015),
             curvature = -0.2, 
             size = .15,
             arrow = arrow(length = unit(0.3, "cm"))) +
  annotate("text", 
           x = (alpha2_w + alpha2_s)/2,
           y=.0022, 
           label=paste0("theta[diff]", "~and~", "delta[diff]"), size = 4, parse=TRUE) +
  scale_color_manual(breaks = c(word_lbl, sent_lbl), values = ggthemes::colorblind_pal()(2)) +
  scale_fill_manual(breaks = c(word_lbl, sent_lbl), values = ggthemes::colorblind_pal()(2)) +
  scale_x_continuous(labels = scales::comma, 
                     limits = c(50, 1000), 
                     breaks = seq(0, 1500, 200)) +
  labs(x = "Interkey intervals [in ms]", 
       colour = "Text location",
       fill = "Text location") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

```
</div>

<div style="float: right; width: 50%;">
\
\
\
\

$$
\text{iki}_{i} \sim\text{ } \theta_\text{location[i]} \times \text{log}\mathcal{N}(\beta + \delta_\text{location[i]}, \sigma_{e'_\text{location[i]}}^2) + \\
   (1 - \theta_\text{location[i]}) \times \text{log}\mathcal{N}(\beta, \sigma_{e_\text{location[i]}}^2)
$$